---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "1.1&nbsp; Introduction to Forecasting"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
options(digits=4, width=55)
```

# Background

## Resources

\large

 * Slides
 * Exercises
 * Textbook
 * Useful links

\begin{alertblock}{}{\centerline{\Large\textbf{\url{robjhyndman.com/nyc2018}}}}
\end{alertblock}

## Brief bio

  - Professor of Statistics, Monash University
  - Editor-in-Chief, \textit{International Journal of Forecasting}

\pause\vspace*{0.3cm}

\begin{block}{How my forecasting methodology is used:}
\begin{itemize}
 \item Pharmaceutical Benefits \rlap{Scheme}
 \item Cancer incidence and mortality
 \item Electricity demand
 \item Ageing population
 \item Fertilizer sales
\end{itemize}
\end{block}

\only<3>{\begin{textblock}{5.5}(6.9,8.1)\begin{alertblock}{}\Large robjhyndman.com\end{alertblock}\end{textblock}}

\vspace*{10cm}

## Assumptions

  * This is not an introduction to R. I assume you are broadly comfortable with R code and the RStudio environment.
  * This is not a statistics course. I assume you are familiar with concepts such as the mean, standard deviation, quantiles, regression, normal distribution, likelihood, etc.
  * This is not a theory course. I am not going to derive anything. I will teach you forecasting tools, when to use them and how to use them most effectively.

## Key reference

\large

\begin{block}{}\bf
\hangafter=1\hangindent=.3cm
 {Hyndman, R.~J. \& Athanasopoulos, G. (2018) \emph{Forecasting: principles and practice}, 2nd ed.}
\end{block}\pause
\begin{alertblock}{}\Large
\centerline{\bf OTexts.org/fpp2/}
\end{alertblock}

\pause

  * Free and online
  * Data sets in associated R package
  * R code for examples

## Poll: How experienced are you in forecasting

  1. Guru: I wrote the book, done it for decades, now I do the conference circuit.
  2. Expert: It has been my full time job for more than a decade.
  3. Skilled: I have been doing it for years.
  1. Comfortable: I understand it and have done it.
  1. Learner: I am still learning.
  1. Beginner: I have heard of it and would like to learn more.
  1. Unknown: What is forecasting? Is that what the weather people do?

## Poll: How proficient are you in using R?

\fontsize{14}{15}\sf

  1. Guru: The R core team come to me for advice.
  1. Expert: I have written several packages on CRAN.
  1. Skilled: I use it regularly and it is an important part of my job.
  1. Comfortable: I use it often and am comfortable with the tool.
  1. User: I use it sometimes, but I am often searching around for the right function.
  1. Learner: I have used it a few times.
  1. Beginner: I've managed to download and install it.
  1. Unknown: Why are you speaking like a pirate?

## Install required packages

```r
install.packages("fpp2", dependencies=TRUE)
```

<!-- ## Getting help with R -->

<!-- \small -->

<!-- ``` -->
<!-- # Search for terms -->
<!-- ??forecasting -->

<!-- # Detailed help -->
<!-- ?forecast -->

<!-- # Worked examples -->
<!-- example("forecast.ar") -->

<!-- # Similar names -->
<!-- apropos("forecast") -->

<!-- # Help on package -->
<!-- help(package="fpp") -->
<!-- ``` -->

## Approximate outline
\vspace*{-0.1cm}\centering\fontsize{12}{12}\sf
\begin{tabular}{rp{6.6cm}l}
  \bf Day & \bf Topic                     & \bf Chapter \\
  \midrule
  1.1     & Time series graphics          & 2 \\
  1.2     & Benchmark methods             & 3 \\
  1.3     & Forecast evaluation           & 3 \\
  1.4     & Exponential smoothing         & 7 \\
  1.5     & ETS state space models        & 7 \\
  \midrule
  2.1     & Transformations               & 3 \\
  2.2     & Seasonality and trends        & 6 \\
  2.3     & Stationarity and differencing & 8 \\
  2.4     & Non-seasonal ARIMA models     & 8 \\
  2.5     & Seasonal ARIMA models         & 8 \\
  \midrule
  3.1     & Dynamic regression            & 9 \\
  3.2     & Multiple seasonality          & 11 \\
  3.3     & Hierarchical forecasting      & 10 \\
  3.4     & Some practical issues         & 12
\end{tabular}

# Case studies

## CASE STUDY 1: Paperware company

\fontsize{12}{14}\sf

\begin{textblock}{7.6}(0.2,1.4)
\textbf{Problem:} Want forecasts of each of hundreds of
items. Series can be stationary, trended or seasonal. They currently
have a large forecasting program written in-house but it doesn't seem
to produce sensible forecasts. They want me to tell them what is
wrong and fix it.

\vspace*{-0.2cm}

\textbf{Additional information}\vspace*{-0.3cm}\fontsize{12}{13}\sf
\begin{itemize}\itemsep=0cm\parskip=0cm
\item  Program  written in COBOL making numerical calculations limited. It is not possible to do any optimisation.
\item Their programmer has little experience in numerical computing.
\item They employ no statisticians and want the program to produce forecasts \rlap{automatically.}
\end{itemize}
\end{textblock}

\placefig{8}{1.4}{width=4.8cm}{tableware2}

## CASE STUDY 1: Paperware company

### Methods currently used

A
: 12 month average

C
: 6 month average

E
: straight line regression over last 12 months

G
: straight line regression over last 6 months

H
: average slope between last year's and this year's values.
  (Equivalent to differencing at lag 12 and taking mean.)

I
: Same as H except over 6 months.

K
: I couldn't understand the explanation.

## CASE STUDY 2: PBS

\full{pills}

## CASE STUDY 2: PBS
\begin{block}{}
The \alert{Pharmaceutical Benefits Scheme} (PBS) is the Australian government drugs subsidy scheme.
\end{block}
\pause

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.

## CASE STUDY 2: PBS

\full{pbs2}

## CASE STUDY 2: PBS

  * In 2001: \$4.5 billion budget, under-forecasted by \$800~million.
  * Thousands of products. Seasonal demand.
  * Subject to covert marketing, volatile products, uncontrollable expenditure.
  * Although monthly data available for 10 years, data are aggregated to annual values, and only the first three years are used in estimating the forecasts.
  * All forecasts being done with the \texttt{FORECAST} function in MS-Excel!

## \large CASE STUDY 3: Airline

\full{ansettlogo}

## CASE STUDY 3: Airline

```{r, echo=FALSE}
autoplot(melsyd[,"Economy.Class"],
  main="Economy class passengers: Melbourne-Sydney",
  xlab="Year",ylab="Thousands")
```

\only<2>{\begin{textblock}{4.2}(7,6)
\begin{alertblock}{}
Not the real data! Or is it?
\end{alertblock}
\end{textblock}}

## CASE STUDY 3: Airline

\textbf{Problem:} how to forecast passenger traffic on major routes?

### Additional information

  * They can provide a large amount of data on previous routes.
  * Traffic is affected by school holidays, special events such as
the Grand Prix, advertising campaigns, competition behaviour, etc.
  * They have a highly capable team of people who are able to do
most of the computing.

# The statistical forecasting perspective

## Sample futures

```{r austa1, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.width=9, fig.height=6}
fit <- ets(austa)
df <- cbind(austa, simulate(fit,10))
for(i in seq(9))
  df <- cbind(df, simulate(fit,10))
colnames(df) <- c("Data", paste("Future",1:10))
autoplot(df) +
  ylim(min(austa),10) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Total international visitors to Australia") +
 scale_colour_manual(values=c('#000000',rainbow(10)),
                     breaks=c("Data",paste("Future",1:10)),
                     name=" ") +
  ylim(.85,10.0)
```

## Forecast intervals

```{r austa2, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.width=8.6, fig.height=6}
autoplot(forecast(fit)) +
  ylab("Millions of visitors") + xlab("Year") +
  ggtitle("Forecasts of total international visitors to Australia") +
  ylim(0.85,10.0)
```

# What can we forecast?

## Forecasting is difficult

\full{hopecasts2}

## What can we forecast?

\full{nasdaq-stock-market}

## What can we forecast?

\full{Forex2}

## What can we forecast?

\full{pills}

## What can we forecast?

\full{elecwires2}

## What can we forecast?

\full{AusBOM}

## What can we forecast?

\full{ts22015}

## What can we forecast?

\full{comet}

## Which is easiest to forecast?

 1. daily electricity demand in 3 days time
 2. timing of next Halley's comet appearance
 3. time of sunrise this day next year
 4. Google stock price tomorrow
 5. Google stock price in 6 months time
 6. maximum temperature tomorrow
 7. exchange rate of \$US/AUS next week
 8. total sales of drugs in Australian pharmacies next month

\pause

 - how do we measure "easiest"?
 - what makes something easy/difficult to forecast?

## Factors affecting forecastability

Something is easier to forecast if:

 - we have a good understanding of the factors that contribute to it
 - there is lots of data available;
 - the forecasts cannot affect the thing we are trying to forecast.
 - there is relatively low natural/unexplainable random variation.
 - the future is somewhat similar to the past

