---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "1.2&nbsp; Time series graphics"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
source("nicefigs.R")
options(digits=4, width=55)
```

# Time series in R

## `ts` objects and `ts` function

\fontsize{13}{13}\sf

A time series is stored in a `ts` object in R:\vspace*{-0.3cm}

 - a list of numbers
 - information about times those numbers were recorded.

### Example

```{r tstable, echo=FALSE, cache=TRUE}
x <- c(123,39,78,52,110)
yr <- 2012:2016
knitr::kable(data.frame(Year=yr,Observation=x), booktabs=TRUE)
```

```r
y <- ts(c(123,39,78,52,110), start=2012)
```

## `ts` objects and `ts` function

For observations that are more frequent than once per year, add a `frequency` argument.

E.g., monthly data stored as a numerical vector `z`:

```r
y <- ts(z, frequency=12, start=c(2003, 1))
```

## `ts` objects and `ts` function

### `ts(data, frequency, start)`

\begin{tabular}{lrl}
\bf Type of data & \hspace*{1.95cm}\bf frequency                 & \bf start example\hspace*{0.25cm} \\
\midrule
Annual           & \only<2->{1}                                  & \only<3->{1995}\\
Quarterly        & \only<4->{4}                                  & \only<5->{c(1995,2)}\\
Monthly          & \only<6->{12}                                 & \only<7->{c(1995,9)}\\
Daily            & \only<8->{7 \emph{or} 365.25}                 & \only<9->{1 \emph{or} c(1995,234)} \\
Weekly           & \only<10->{52.18}                             & \only<11->{c(1995,23)}\\
Hourly           & \only<12->{24 \emph{or} 168 \emph{or} 8,766}  & \only<13->{1}\\
Half-hourly      & \only<14->{48 \emph{or} 336 \emph{or} 17,532} & \only<15>{1}
\end{tabular}

## Australian GDP
\fontsize{13}{15}\sf

```r
ausgdp <- ts(x, frequency=4, start=c(1971,3))
```

  * Class: "ts"
  * Print and plotting methods available.

```{r, echo=TRUE}
ausgdp
```

## Australian GDP

```{r, echo=TRUE, fig.height=4}
autoplot(ausgdp)
```

## Residential electricity sales

```{r}
elecsales
```

## Time plots

\small

```{r, echo=TRUE, fig.height=4}
autoplot(melsyd[,"Economy.Class"])
```

## Time plots

\small

```{r a10, echo=TRUE}
autoplot(a10) + ylab("$ million") + xlab("Year") +
         ggtitle("Antidiabetic drug sales")
```

# Lab session 1

##
\fontsize{48}{60}\sf\centering
**Lab Session 1**

# Seasonal plots

## Seasonal plots

\small

```{r, echo=TRUE, fig.height=3.5}
ggseasonplot(a10, ylab="$ million",
  year.labels=TRUE, year.labels.left=TRUE) +
  ggtitle("Seasonal plot: antidiabetic drug sales")
```

## Seasonal plots

  * Data plotted against the individual "seasons" in which the data were observed.  (In this case a "season" is a month.)
  * Something like a time plot except that the data from each season are overlapped.
  * Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
  * In R: `ggseasonplot`

## Seasonal polar plots

\small

```{r, fig.height=5, out.width="7.5cm"}
ggseasonplot(a10, polar=TRUE) + ylab("$ million")
```

## Seasonal subseries plots

\fontsize{10}{12}\sf

```{r, echo=TRUE, fig.height=3.5}
ggsubseriesplot(a10) + ylab("$ million") +
  ggtitle("Seasonal subseries plot: antidiabetic drug sales")
```

## Seasonal subseries plots

  * Data for each season collected together in time plot as separate time series.
  * Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be visualized.
  * In R: `ggsubseriesplot`

## Quarterly Australian Beer Production

```{r}
beer <- window(ausbeer,start=1992)
autoplot(beer)
```

## Quarterly Australian Beer Production

```{r}
ggseasonplot(beer,year.labels=TRUE)
```

## Quarterly Australian Beer Production

```{r}
ggsubseriesplot(beer)
```

# Lab session 2
##
\fontsize{48}{60}\sf\centering
**Lab Session 2**

# Seasonal or cyclic?

## Time series patterns

Trend
  : pattern exists when there is a long-term increase or decrease in the data.

Seasonal
  : pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic
  : pattern exists when data exhibit rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).

## Time series components

### Differences between seasonal and cyclic patterns:

* seasonal pattern constant length; cyclic pattern variable length
* average length of cycle longer than length of seasonal pattern
* magnitude of cycle more variable than magnitude of seasonal pattern

## Time series patterns
\small

```{r, fig.height=3.9}
autoplot(window(elec, start=1980)) +
  ggtitle("Australian electricity production") +
  xlab("Year") + ylab("GWh")
```

## Time series patterns
\small

```{r, fig.height=3.9}
autoplot(bricksq) +
  ggtitle("Australian clay brick production") +
  xlab("Year") + ylab("million units")
```

## Time series patterns
\small

```{r, fig.height=3.9}
autoplot(hsales) +
  ggtitle("Sales of new one-family houses, USA") +
  xlab("Year") + ylab("Total sales")
```

## Time series patterns
\small

```{r, fig.height=3.9}
autoplot(ustreas) +
  ggtitle("US Treasury Bill Contracts") +
  xlab("Day") + ylab("price")
```

## Time series patterns
\small

```{r, fig.height=3.9}
autoplot(lynx) +
  ggtitle("Annual Canadian Lynx Trappings") +
  xlab("Year") + ylab("Number trapped")
```

## Seasonal or cyclic?

\alert{Differences between seasonal and cyclic patterns:}

  * seasonal pattern constant length; cyclic pattern variable length
  * average length of cycle longer than length of seasonal pattern
  * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

\begin{alertblock}{}
The timing of peaks and troughs is predictable with seasonal data, but unpredictable in the long term with cyclic data.
\end{alertblock}

# Lag plots and autocorrelation

## Example: Beer production

```r
beer <- window(ausbeer, start=1992)
gglagplot(beer, lags=9, do.lines=FALSE,
  continuous=FALSE)
```

## Example: Beer production

```{r, echo=FALSE, fig.height=6, fig.width=6, out.width="8cm"}
beer <- window(ausbeer, start=1992)
gglagplot(beer, lags=9, do.lines=FALSE, continuous=FALSE)
```

## Lagged scatterplots

  * Each graph shows $y_t$ plotted against $y_{t-k}$ for
different values of $k$.
  * The autocorrelations are the correlations associated
with these scatterplots.
  * ACF (autocorrelation function):
     * $r_1=\text{Correlation}(y_{t}, y_{t-1})$
     * $r_2=\text{Correlation}(y_{t}, y_{t-2})$
     * $r_3=\text{Correlation}(y_{t}, y_{t-3})$
     * etc.
  * If there is **seasonality**, the ACF at the seasonal lag (e.g., 12 for monthly data) will be **large and positive**.

## Autocorrelation

\small
Results for first 9 lags for beer data:

\footnotesize

```{r, echo=FALSE}
beeracf <- matrix(acf(c(beer), lag.max=9,
                      plot=FALSE)$acf[-1,,1], nrow=1)
colnames(beeracf) <- paste("$r_",1:9,"$",sep="")
knitr::kable(beeracf, booktabs=TRUE,
             align="c", digits=3,
             format.args=list(nsmall=3))
```

```{r beeracf, fig.height=2.5}
ggAcf(beer)
```

## Aus monthly electricity production

```{r, fig.height=4}
elec2 <- window(elec, start=1980)
autoplot(elec2)
```

## Aus monthly electricity production

```{r}
ggAcf(elec2, lag.max=48)
```

## Google stock price

```{r}
autoplot(goog)
```

## Google stock price

```{r}
ggAcf(goog, lag.max=100)
```

## Which is which?

```{r, fig.height=6, fig.width=12, echo=FALSE, out.width="11.7cm"}
tp1 <- autoplot(cowtemp) + xlab("") + ylab("chirps per minute") +
  ggtitle("1. Daily temperature of cow")
tp2 <- autoplot(USAccDeaths/1e3) + xlab("") + ylab("thousands") +
  ggtitle("2. Monthly accidental deaths")
tp3 <- autoplot(AirPassengers) + xlab("") + ylab("thousands") +
  ggtitle("3. Monthly air passengers")
tp4 <- autoplot(mink/1e3) + xlab("") + ylab("thousands") +
  ggtitle("4. Annual mink trappings")
acfb <- ggAcf(cowtemp, ci=0) + xlab("") + ggtitle("B") + ylim(-0.4,1)
acfa <- ggAcf(USAccDeaths, ci=0) + xlab("") + ggtitle("A") + ylim(-0.4,1)
acfd <- ggAcf(AirPassengers, ci=0) + xlab("") + ggtitle("D") + ylim(-0.4,1)
acfc <- ggAcf(mink, ci=0) + xlab("") + ggtitle("C") + ylim(-0.4,1)
gridExtra::grid.arrange(tp1,tp2,tp3,tp4,
                        acfa,acfb,acfc,acfd,nrow=2)
```

# White noise

## Example: White noise

```{r, fig.height=3.5}
wn <- ts(rnorm(36))
autoplot(wn)
```

\only<2>{
\begin{textblock}{10}(1.4,7)
\begin{alertblock}{}
White noise data is uncorrelated across time with zero mean and constant variance.

(Technically, we require independence as well.)
\end{alertblock}
\end{textblock}}

## Example: White noise

```{r, results='asis', echo=FALSE}
wnacf <- matrix(acf(c(wn), lag.max=10,
                      plot=FALSE)$acf[-1,,1], nrow=1)
colnames(wnacf) <- paste("$r_{",1:10,"}$",sep="")
print(xtable::xtable(t(wnacf)),
    sanitize.rownames.function=identity,
    booktabs=TRUE,
    include.colnames = FALSE,
    hline.after = FALSE,
    size='footnotesize',
    comment=FALSE,
    floating=FALSE)
```

```{r, echo=FALSE}
savepdf("wnacf")
ggAcf(wn, max.lag=15)
endpdf()
```

\placefig{4.5}{1.6}{width=8cm}{wnacf}
\vspace*{0.3cm}

 * Sample autocorrelations for white noise series.
 * Expect each autocorrelation to be close to zero.
 * Blue lines show 95% critical values.

## Example: Pigs slaughtered
\fontsize{12}{13}\sf

```{r, fig.height=3.5}
pigs2 <- window(pigs, start=1990)
autoplot(pigs2) +
  xlab("Year") + ylab("thousands") +
  ggtitle("Number of pigs slaughtered in Victoria")
```

## Example: Pigs slaughtered

```{r}
ggAcf(pigs2)
```

# Lab session 3
##
\fontsize{48}{60}\sf\centering
**Lab Session 3**
