---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "2.1&nbsp; Transformations"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
source("nicefigs.R")
options(digits=4, width=55)
```

# Variance stabilization

## Variance stabilization
\fontsize{13}{14}\sf

If the data show different variation at different levels of the series, then a transformation can be useful.
\pause

Denote original observations as $y_1,\dots,y_n$ and transformed
observations as $w_1, \dots, w_n$.
\pause

\begin{block}{Mathematical transformations for stabilizing
variation}
\begin{tabular}{llc}
Square root & $w_t = \sqrt{y_t}$ & $\downarrow$ \\[0.2cm]
Cube root & $w_t = \sqrt[3]{y_t}$ & Increasing \\[0.2cm]
Logarithm & $w_t = \log(y_t)$  & strength
\end{tabular}
\end{block}
\pause

Logarithms, in particular, are useful because they are more interpretable: changes in a log value are **relative (percent) changes on the original scale**.

## Variance stabilization

```{r, echo=FALSE, fig.height=4.6}
autoplot(elec) +
  xlab("Year") + ylab("") +
  ggtitle("Electricity production")
```

## Variance stabilization

```{r, echo=FALSE, fig.height=4.6}
autoplot(elec^0.5) +
  xlab("Year") + ylab("") +
  ggtitle("Square root electricity production")
```

## Variance stabilization

```{r, echo=FALSE, fig.height=4.6}
autoplot(elec^0.33333) +
  xlab("Year") + ylab("") +
  ggtitle("Cube root electricity production")
```

## Variance stabilization

```{r, echo=FALSE, fig.height=4.6}
autoplot(log(elec)) +
  xlab("Year") + ylab("") +
  ggtitle("Log electricity production")
```

## Variance stabilization

```{r, echo=FALSE, fig.height=4.6}
autoplot(-1/elec) +
  xlab("Year") + ylab("") +
  ggtitle("Inverse electricity production")
```

# Box-Cox transformations

## Box-Cox transformations

Each of these transformations is close to a member of the
family of \textbf{Box-Cox transformations}:
$$w_t = \left\{\begin{array}{ll}
        \log(y_t),      & \quad \lambda = 0; \\
        (y_t^\lambda-1)/\lambda ,         & \quad \lambda \ne 0.
\end{array}\right.
$$\pause

* $\lambda=1$: (No substantive transformation)
* $\lambda=\frac12$: (Square root plus linear transformation)
* $\lambda=0$: (Natural logarithm)
* $\lambda=-1$: (Inverse plus 1)

## Box-Cox transformations


```{r elec5, cache=TRUE, echo=FALSE}
library(latex2exp)
lambda <- seq(1, -1, by=-0.01)
for(i in seq_along(lambda))
{
  savepdf(paste("elecBC",i,sep=""))
  print(autoplot(BoxCox(elec,lambda[i])) + xlab("Year") +
    ylab("") +
    ggtitle(
      TeX(paste("Transformed Australian electricity demand:  $\\lambda =",format(lambda[i],digits=2,nsmall=2),"$"))
    ) +
    scale_y_continuous(breaks=NULL,minor_breaks=NULL) +
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()))
  endpdf()
}
```

\vspace*{0.2cm}
\centerline{\animategraphics[controls,buttonsize=0.3cm,width=12.5cm]{4}{elecBC}{1}{201}}

## Box-Cox transformations

```{r elec6,echo=TRUE,fig.height=4}
autoplot(BoxCox(elec,lambda=1/3))
```

## Box-Cox transformations

* $y_t^\lambda$ for $\lambda$ close to zero behaves like logs.
* If some $y_t=0$, then must have $\lambda>0$
* if some $y_t<0$, no power transformation is possible unless all $y_t$ adjusted by **adding a constant to all values**.
* Simple values of $\lambda$ are easier to explain.
* Results are  relatively insensitive to  $\lambda$.
* Often no transformation ($\lambda=1$) needed.
* Transformation can have very large effect on PI.
* Choosing $\lambda=0$ is a simple way to force forecasts to be positive

## Automated Box-Cox transformations

```{r, echo=TRUE}
(BoxCox.lambda(elec))
```
\pause

* This attempts to balance the seasonal fluctuations and random variation across the series.
* Always check the results.
* A low value of $\lambda$ can give extremely large prediction intervals.

# Back-transformation

## Back-transformation

We must reverse the transformation (or \textit{back-transform}) to obtain
forecasts on the original scale.  The reverse Box-Cox transformations are given
by
$$ y_t = \left\{\begin{array}{ll}
        \exp(w_t),      & \quad \lambda = 0; \\
        (\lambda W_t+1)^{1/\lambda} ,   & \quad \lambda \ne 0.
\end{array}\right.$$

## Back-transformation

```{r,echo=TRUE,fig.height=3.6}
fit <- snaive(elec, lambda=1/3)
autoplot(fit)
```

## Back-transformation

```{r,echo=TRUE,fig.height=4}
autoplot(fit, include=120)
```

## Back transformation

  * Back-transformed point forecasts are medians.
  * Back-transformed PI have the correct coverage.

\pause

**Back-transformed means**

Let $X$ be have mean $\mu$ and variance $\sigma^2$.

Let $f(x)$ be back-transformation function, and $Y=f(X)$.

\begin{alertblock}{}
\centerline{$\E[Y] = \E[f(X)] = f(\mu) + \frac12 \sigma^2[f''(\mu)]^2.$}
\end{alertblock}

## Back transformation
\vspace*{-0.2cm}

**Box-Cox back-transformation:**\vspace*{-0.0cm}
\begin{align*}
y_t &= \left\{\begin{array}{ll}
        \exp(w_t)      & \quad \lambda = 0; \\
        (\lambda W_t+1)^{1/\lambda}  & \quad \lambda \ne 0.
\end{array}\right. \\
f(x) &= \begin{cases}
                        e^x & \quad\lambda=0;\\
 (\lambda x + 1)^{1/\lambda} & \quad\lambda\ne0.
 \end{cases}\\
f''(x) &= \begin{cases}
                        e^x & \quad\lambda=0;\\
 (1-\lambda)(\lambda x + 1)^{1/\lambda-2} & \quad\lambda\ne0.
 \end{cases}
\end{align*}\pause
\begin{alertblock}{}
\centerline{$\E[Y] = \begin{cases}
                        e^\mu\left[1+\frac{\sigma^2}{2}\right] & \quad\lambda=0;\\
 (\lambda \mu + 1)^{1/\lambda}\left[1+\frac{\sigma^2(1-\lambda)}{2(\lambda\mu+1)^2}\right] & \quad\lambda\ne0.
 \end{cases}$}
\end{alertblock}


## Back-transformation
\fontsize{13}{13}\sf

```{r,echo=TRUE,fig.height=3.7}
elec %>% snaive(lambda=1/3, biasadj=FALSE) %>%
  autoplot(include=120)
```

## Back-transformation
\fontsize{13}{13}\sf

```{r,echo=TRUE,fig.height=3.7}
elec %>% snaive(lambda=1/3, biasadj=TRUE) %>%
  autoplot(include=120)
```

## Back-transformation
\fontsize{13}{13}\sf

```{r,echo=TRUE,fig.height=3.7}
eggs %>% ses(lambda=1/3, biasadj=FALSE) %>%
  autoplot
```

## Back-transformation
\fontsize{13}{13}\sf

```{r,echo=TRUE,fig.height=3.7}
eggs %>% ses(lambda=1/3, biasadj=TRUE) %>%
  autoplot
```

# Lab session 11
##

\fontsize{48}{60}\sf\centering
**Lab Session 11**



