---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "2.2&nbsp; Seasonality and trends"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
options(digits=4, width=55)
```

# Time series components

## Time series patterns

Trend
  : pattern exists when there is a long-term increase or decrease in the data.

Seasonal
  : pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic
  : pattern exists when data exhibit rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).

## Time series decomposition

\begin{block}{}\vspace*{-0.3cm}
\[ Y_t = S_t + T_t + R_t \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $Y_t=$ & data at period $t$ \\
      & $S_t=$ & seasonal component at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $R_t=$ & remainder (or irregular or error) component at period $t$
\end{tabular}\pause

* Additive model appropriate if magnitude of seasonal fluctuations does not vary with level.
* If seasonal are proportional to level of series, use a Box-Cox transformation first.


## Euro electrical equipment

```{r mstl, fig.height=4}
fit <- mstl(elecequip)
autoplot(fit)
```

## Euro electrical equipment
\fontsize{10}{11}\sf

```{r mstl3, dependson='mstl', fig.height=3.5}
autoplot(elecequip, series="Data") +
 autolayer(trendcycle(fit), series="Trend") +
 ylab("New orders index") + xlab("") +
 ggtitle("Electrical equipment manufacturing (Euro area)")
```

## Euro electrical equipment
\fontsize{13}{15}\sf

```{r mstl2, dependson='mstl'}
ggmonthplot(seasonal(fit)) + ylab("Seasonal")
```

## Seasonal adjustment

Seasonally adjusted data given by $Y_t - S_t = T_t + E_t$

\fontsize{11}{11}\sf

```{r fig.height=3.3, dependson='mstl'}
autoplot(elecequip, series="Data") +
  autolayer(seasadj(fit), series="SeasAdjust") +
  ylab("New orders index") +
  ggtitle("Electrical equipment manufacturing")
```

## History of time series decomposition

* Classical method originated in 1920s.
* Census II method introduced in 1957. Basis for modern X-12-ARIMA method.
* STL method introduced in 1983
* TRAMO/SEATS introduced in 1990s.

# STL decomposition

## STL decomposition
\fontsize{14}{15}\sf

* STL: "Seasonal and Trend decomposition using Loess",
* Very versatile and robust.
* Will handle any type of seasonality.
* Seasonal component allowed to change over time, and rate of change controlled by user.
* Smoothness of trend-cycle also controlled by user.
* Robust to outliers
* No trading day or calendar adjustments.
* Only additive.

## Euro electrical equipment
\fontsize{10}{12}\sf

```{r, echo=TRUE, fig.height=4}
elecequip %>%
  mstl(s.window=5) %>%
  autoplot()
```

## Euro electrical equipment
\fontsize{10}{12}\sf

```{r, echo=TRUE, fig.height=4}
elecequip %>%
  mstl(t.window=15, s.window='periodic', robust=TRUE) %>%
  autoplot()
```

## STL decomposition in R

* `t.window` controls wiggliness of trend component.
* `s.window` controls variation on seasonal component.
* `seasonal()` extracts seasonal component
* `trendcycle()` extracts trend component
* `remainder()` extracts remainder component
* `seasadj()` computes seasonally adjusted data



# Lab session 12
##
\fontsize{48}{60}\sf\centering
**Lab Session 12**

# Forecasting and decomposition

## Forecasting and decomposition

* Forecast seasonal component by repeating the last year (`snaive`)
* Forecast seasonally adjusted data using non-seasonal time series method. E.g.,
    - Holt's method
    - Random walk with drift model
* Combine forecasts of seasonal component with forecasts of seasonally adjusted data to get forecasts of original data.
* Sometimes a decomposition is useful just for understanding the data before building a separate forecasting model.

## Seas adj elec equipment
\fontsize{11}{12}\sf

```{r, fig.height=3.5}
mstl(elecequip, t.window=15, s.window="periodic") %>%
  seasadj() %>% naive(h=24) %>%
  autoplot() + ylab("New orders index") +
  ggtitle("Naive forecasts of seasonally adjusted data")
```

## Seas adj elec equipment
\fontsize{11}{12}\sf

```{r, fig.height=3.5}
mstl(elecequip, t.window=15, s.window="periodic") %>%
  forecast(method="naive", h=24) %>%
  autoplot() + ylab("New orders index") +
  ggtitle("Reseasonalized naive forecasts")
```

## Decomposition and prediction intervals

* It is common to take the prediction intervals from the seasonally adjusted forecasts and modify them with the seasonal component.
* This ignores the uncertainty in the seasonal component estimate.
* It also ignores the uncertainty in the future seasonal pattern.

## Some more R functions

```r
fcast <- stlf(elecequip, method='naive')

fcast <- stlf(elecequip, method='naive',
  h=36, s.window=11, robust=TRUE)
```

# Lab session 13
##
\fontsize{48}{60}\sf\centering
**Lab Session 13**

# Lab session 14
##
\fontsize{48}{60}\sf\centering
**Lab Session 14**

