---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "2.3&nbsp; Stationarity and differencing"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
options(digits=4, width=55)
```

# Stationarity

## Stationarity

\begin{block}{Definition}
If $\{y_t\}$ is a stationary time series, then for all $s$, the distribution of $(y_t,\dots,y_{t+s})$ does not depend on $t$.
\end{block}\pause

A \textbf{stationary series} is:

*  roughly horizontal
*  constant variance
*  no patterns predictable in the long-term

## Stationary?

```{r, echo=FALSE}
autoplot(dj) + ylab("Dow Jones Index") + xlab("Day")
```

## Stationary?

```{r, echo=FALSE}
autoplot(diff(dj)) + ylab("Change in Dow Jones Index") + xlab("Day")
```

## Stationary?

```{r, echo=FALSE}
autoplot(strikes) + ylab("Number of strikes") + xlab("Year")
```

## Stationary?

```{r, echo=FALSE}
autoplot(hsales) + xlab("Year") + ylab("Total sales") +
  ggtitle("Sales of new one-family houses, USA")
```

## Stationary?
```{r, echo=FALSE}
autoplot(eggs) + xlab("Year") + ylab("$") +
  ggtitle("Price of a dozen eggs in 1993 dollars")
```

## Stationary?
```{r, echo=FALSE}
autoplot(window(pigs/1e3, start=1990)) + xlab("Year") + ylab("thousands") +
  ggtitle("Number of pigs slaughtered in Victoria")
```

## Stationary?

```{r, echo=FALSE}
autoplot(lynx) + xlab("Year") + ylab("Number trapped") +
  ggtitle("Annual Canadian Lynx Trappings")
```

## Stationary?

```{r, echo=FALSE}
autoplot(window(ausbeer, start=1992)) + xlab("Year") + ylab("megalitres") +
  ggtitle("Australian quarterly beer production")
```

## Stationarity

\begin{block}{Definition}
If $\{y_t\}$ is a stationary time series, then for all $s$, the distribution of $(y_t,\dots,y_{t+s})$ does not depend on $t$.
\end{block}\pause\vspace*{0.4cm}

Transformations help to \textbf{stabilize the variance}.

For ARIMA modelling, we also need to \textbf{stabilize the mean}.

## Non-stationarity in the mean
\alert{Identifying non-stationary series}

* time plot.

* The ACF of stationary data drops to zero relatively quickly
* The ACF of non-stationary data decreases slowly.
* For non-stationary data, the value of $r_1$ is often
     large and positive.

## Example: Dow-Jones index
\fontsize{11}{11}\sf

```{r}
autoplot(dj) + ylab("Dow Jones Index") + xlab("Day")
```

## Example: Dow-Jones index
\fontsize{11}{11}\sf

```{r}
ggAcf(dj)
```

## Example: Dow-Jones index
\fontsize{11}{11}\sf

```{r}
autoplot(diff(dj)) +
  ylab("Change in Dow Jones Index") + xlab("Day")
```

## Example: Dow-Jones index
\fontsize{11}{11}\sf

```{r}
ggAcf(diff(dj))
```

# Differencing

## Differencing

* Differencing helps to \textbf{stabilize the mean}.
* The differenced series is the \emph{change} between each observation
in the original series: ${y'_t = y_t - y_{t-1}}$.
* The differenced series will have only $T-1$ values since it is not possible to calculate a difference $y_1'$ for the first observation.

## Second-order  differencing

Occasionally the differenced data will not appear stationary and it
may be necessary to difference the data a second time:\pause
\begin{align*}
y''_{t}  &=  y'_{t}  - y'_{t - 1} \\
&= (y_t - y_{t-1}) - (y_{t-1}-y_{t-2})\\
&= y_t - 2y_{t-1} +y_{t-2}.
\end{align*}\pause\vspace*{-0.75cm}

* $y_t''$ will have  $T-2$  values.
* In practice,  it is almost never necessary to go beyond second-order
differences.

## Seasonal differencing

A seasonal difference is the difference between an observation and the corresponding observation from the previous year.\pause
$${y'_t = y_t - y_{t-m}}$$
where $m=$ number of seasons.\pause

* For monthly data $m=12$.
* For quarterly data $m=4$.

## Electricity production
```{r, echo=TRUE, fig.height=4}
usmelec %>% autoplot()
```

## Electricity production
```{r, echo=TRUE, fig.height=4}
usmelec %>% log() %>% autoplot()
```

## Electricity production

```{r, echo=TRUE, fig.height=3.5}
usmelec %>% log() %>% diff(lag=12) %>%
  autoplot()
```

## Electricity production
```{r, echo=TRUE, fig.height=3.5}
usmelec %>% log() %>% diff(lag=12) %>%
  diff(lag=1) %>% autoplot()
```

## Electricity production

* Seasonally differenced series is closer to being stationary.
* Remaining non-stationarity  can be removed with further first difference.

If $y'_t = y_t - y_{t-12}$ denotes seasonally differenced series, then twice-differenced series is\vspace*{-0.1cm}
\begin{block}{}\vspace*{-0.2cm}
\begin{align*}
y^*_t &= y'_t - y'_{t-1} \\
      &= (y_t - y_{t-12}) - (y_{t-1} - y_{t-13}) \\
      &= y_t - y_{t-1} - y_{t-12} + y_{t-13}\: .
\end{align*}
\end{block}\vspace*{10cm}

## Seasonal differencing
\fontsize{14}{15}\sf

When both seasonal and first differences are applied\dots\pause

* it makes no difference
which is done first---the result will be the same.
* If seasonality is strong, we recommend that seasonal differencing be done first because sometimes the resulting series will be stationary and there will be no need for further first difference.
\pause

###
It is important that if differencing is used, the differences are
interpretable.

## Interpretation of differencing

* first differences are the change between \textbf{one observation and the
next};
* seasonal differences are the change between \textbf{one year to the
next}.
\pause

But taking lag 3 differences for yearly data, for example, results in a model which cannot be sensibly interpreted.

# Unit root tests
## Unit root tests

\alert{Statistical tests to determine the required order of differencing.}

1. Augmented Dickey Fuller test: null hypothesis is that the data are non-stationary and non-seasonal.
2. Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test: null  hypothesis is that the data are stationary and non-seasonal.
3. Other tests available for seasonal data.

## KPSS test
\fontsize{10}{11}\sf

```{r, echo=TRUE}
library(urca)
summary(ur.kpss(goog))
```

\pause

```{r, echo=TRUE}
ndiffs(goog)
```

## Automatically selecting differences

STL decomposition: $y_t = T_t+S_t+R_t$

Seasonal strength $F_s = \max\big(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\big)$

If $F_s > 0.64$, do one seasonal difference.

\pause\fontsize{12}{15}\sf\vspace*{1cm}

```{r, echo=TRUE}
usmelec %>% log() %>% nsdiffs()
usmelec %>% log() %>% diff(lag=12) %>% ndiffs()
```

# Lab session 15
##
\fontsize{48}{60}\sf\centering
**Lab Session 15**

# Backshift notation

## Backshift notation

A  very useful notational device is the backward  shift operator,  $B$,  which is used as follows:
$$
{B y_{t}  =  y_{t - 1}} \: .
$$\pause
In  other  words,   $B$,  operating on  $y_{t}$,   has  the
effect   of   \textbf{shifting  the  data  back  one   period}. \pause
Two applications of  $B$  to  $y_{t}$ \textbf{shifts the data  back  two
periods}:
$$
B(By_{t})  =  B^{2}y_{t}  =  y_{t-2}\: .
$$\pause
For  monthly  data, if we wish to shift attention  to  ``the
same  month last year,''  then  $B^{12}$
is used,  and  the
notation is  $B^{12}y_{t}$  =  $y_{t-12}$.

## Backshift notation

The   backward   shift  operator  is  convenient   for describing  the
process  of  \textit{differencing}. \pause
A first difference can be written as
$$
y'_{t}  =  y_{t} -   y_{t-1} = y_t - By_{t}  =  (1 - B)y_{t}\: .
$$\pause
Note  that a first difference is represented by  $(1 -  B)$.
\pause

Similarly,   if second-order differences (i.e.,   first
differences  of  first differences) have  to  be  computed,
then:
\[
y''_{t}  =  y_{t} -   2y_{t - 1}  +  y_{t - 2} = (1 - B)^{2} y_{t}\: .
\]

## Backshift notation

* Second-order difference  is  denoted   $(1- B)^{2}$.

* \textit{Second-order  difference} is not  the  same  as  a \textit{second  difference},  which would  be  denoted $1- B^{2}$;

* In general,  a  $d$th-order difference can be written as
$$(1 - B)^{d} y_{t}.$$

* A seasonal difference followed by a first difference can be written as
$$ (1-B)(1-B^m)y_t\: .$$

## Backshift notation

The ``backshift'' notation is convenient because the terms can be multiplied
together to see the combined effect.
\begin{align*}
(1-B)(1-B^m)y_t &= (1 - B - B^m + B^{m+1})y_t \\
&= y_t-y_{t-1}-y_{t-m}+y_{t-m-1}.
\end{align*}\pause
For monthly data, $m=12$ and we obtain the same result as earlier.

