---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "3.2&nbsp; Forecasting with multiple seasonality"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
options(digits=4, width=55)
```

# Time series with complex seasonality

## Examples
\fontsize{11}{11}\sf

```{r gasolinedata}
autoplot(gasoline) +
  xlab("Year") + ylab("Thousands of barrels per day") +
  ggtitle("Weekly US finished motor gasoline products")
```

## Examples

```{r callsdata, message=FALSE, warning=FALSE, echo=FALSE,fig.height=5}
p1 <- autoplot(calls) +
  ylab("Call volume") + xlab("Weeks") +
  scale_x_continuous(breaks=seq(1,33,by=2)) +
  ggtitle("5 minute call volume at North American bank")
p2 <- autoplot(window(calls, end=4)) +
  ylab("Call volume") + xlab("Weeks") +
  scale_x_continuous(minor_breaks = seq(1,4,by=0.2))
gridExtra::grid.arrange(p1,p2)
```

## Examples

```{r turk, echo=FALSE}
telec <- read.csv("../data/turkey_elec.csv")
telec <- msts(telec/1e3, start=2000, seasonal.periods = c(7,354.37,365.25))
autoplot(telec) +
  ggtitle("Turkish daily electricity demand") +
  xlab("Year") + ylab("Electricity Demand (GW)")
```


# STL with multiple seasonal periods

## STL with multiple seasonal periods
\fontsize{11}{11}\sf

```{r callsmstl, fig.height=4.6}
calls %>% mstl() %>% autoplot() + xlab("Week")
```

## STL with multiple seasonal periods
\fontsize{11}{11}\sf

```{r callsmstlf}
calls %>%  stlf() %>% autoplot() + xlab("Week")
```

## STL with multiple seasonal periods
\fontsize{11}{11}\sf

```{r callsmstlf2}
calls %>%  stlf() %>% autoplot(include=5*169) + xlab("Week")
```

# Dynamic harmonic regression

## Dynamic harmonic regression
\fontsize{11}{11}\sf

```{r callsharmonics, echo=TRUE, fig.asp=0.45}
fit <- auto.arima(calls, seasonal=FALSE, lambda=0,
         xreg=fourier(calls, K=c(10,10)))
fit %>%
  forecast(xreg=fourier(calls, K=c(10,10), h=2*169)) %>%
  autoplot(include=5*169) +
    ylab("Call volume") + xlab("Weeks")
```

```{r callsorder, echo=FALSE}
ncoef <- length(coef(fit))
arma <- sum(arimaorder(fit)[-2])
nf169 <- sum(grepl("169",names(coef(fit))))
nf845 <- sum(grepl("845",names(coef(fit))))
if(ncoef != (arma+nf169+nf845))
  stop("Coefficients don't add up")
```

This is a large model, containing `r ncoef` parameters: `r arma` ARMA coefficients, `r nf169` Fourier coefficients for frequency 169, and `r nf845` Fourier coefficients for frequency 845. We don't use all the Fourier terms for frequency 845 because there is some overlap with the terms of frequency 169 (since $845=5\times169$).

# TBATS model

## TBATS model

\begin{alertblock}{\Large TBATS}
\textbf{\Large T}rigonometric terms for seasonality\\
\textbf{\Large B}ox-Cox transformations for heterogeneity\\
\textbf{\Large A}RMA errors for short-term dynamics\\
\textbf{\Large T}rend (possibly damped)\\
\textbf{\Large S}easonal (including multiple and non-integer periods)
\end{alertblock}

## TBATS model
\vspace*{-0.8cm}\fontsize{13}{14}\sf

\begin{align*}
y_t&= \text{observation at time $t$}\\
y_t^{(\omega)} &= \begin{cases} (y_t^\omega-1)/\omega & \text{if $\omega\ne0$};\\
\log y_t & \text{if $\omega=0$}.\end{cases}\hspace*{10cm}\\
y_t^{(\omega)} &= \ell_{t-1} + \phi b_{t-1} + \sum_{i=1}^M s_{t-m_i}^{(i)} + d_t\\
\ell_t &= \ell_{t-1} + \phi b_{t-1} + \alpha d_t\\
b_t &= (1-\phi) b + \phi b_{t-1} + \beta d_{t}\\
d_t &= \sum_{i=1}^p \phi_i d_{t-i} + \sum_{j=1}^q \theta_j \varepsilon_{t-j} + \varepsilon_t\\
s_t^{(i)} &= \sum_{j=1}^{k_i} s_{j,t}^{(i)}
\end{align*}
\begin{textblock}{8}(4.5,7.5)
\begin{align*}
s_{j,t}^{(i)} &= \phantom{-}s_{j,t-1}^{(i)}\cos \lambda_j^{(i)} + s_{j,t-1}^{*(i)}\sin \lambda_j^{(i)} + \gamma_1^{(i)} d_t\\
s_{j,t}^{(i)} &= -s_{j,t-1}^{(i)}\sin \lambda_j^{(i)} + s_{j,t-1}^{*(i)}\cos \lambda_j^{(i)} + \gamma_2^{(i)} d_t
\end{align*}
\end{textblock}

\only<2->{\begin{textblock}{5}(7.3,1.8)\begin{block}{}Box-Cox transformation\end{block}\end{textblock}}
\only<3->{\begin{textblock}{5}(7.3,3.4)\begin{block}{}$M$ seasonal periods\end{block}\end{textblock}}
\only<4->{\begin{textblock}{5}(7.3,4.7)\begin{block}{}global and local trend\end{block}\end{textblock}}
\only<5->{\begin{textblock}{5}(7.3,6.0)\begin{block}{}ARMA error\end{block}\end{textblock}}
\only<6->{\begin{textblock}{5}(7.3,6.8)\begin{block}{}Fourier-like seasonal terms\end{block}\end{textblock}}

\only<7>{\begin{textblock}{4}(2.5,2.5)\large\begin{alertblock}{TBATS}
\textbf{T}rigonometric\\
\textbf{B}ox-Cox\\
\textbf{A}RMA\\
\textbf{T}rend\\
\textbf{S}easonal
\end{alertblock}\end{textblock}}

## Complex seasonality
\fontsize{12}{12}\sf

```{r gasoline, echo=TRUE, fig.height=4}
gasoline %>% tbats() %>% forecast() %>% autoplot()
```

## Complex seasonality
\fontsize{12}{12}\sf

```{r callcentref, echo=TRUE, fig.height=4}
calls %>% tbats() %>% forecast() %>% autoplot()
```

## Complex seasonality
\fontsize{12}{12}\sf

```{r telecf, echo=TRUE, fig.height=4}
telec %>% tbats() %>% forecast() %>% autoplot()
```


## TBATS model

\begin{alertblock}{\Large TBATS}\fontsize{13.5}{19}\sf
\textbf{\Large T}rigonometric terms for seasonality\\
\textbf{\Large B}ox-Cox transformations for heterogeneity\\
\textbf{\Large A}RMA errors for short-term dynamics\\
\textbf{\Large T}rend (possibly damped)\\
\textbf{\Large S}easonal (including multiple and non-integer periods)
\end{alertblock}\vspace*{0.1cm}\fontsize{13}{15}\sf

* Handles non-integer seasonality, multiple seasonal periods.
* Entirely automated
* Prediction intervals often too wide
* Very slow on long series


\vspace*{10cm}

# Lab session 21
##
\fontsize{48}{60}\sf\centering
**Lab Session 21**

