---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "3.3&nbsp; Hierarchical forecasting"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
header_includes:
  - \def\bY{\bm{y}}
  - \def\by{\bm{y}}
  - \def\bS{\bm{S}}
  - \def\bI{\text{\rm\textbf{I}}}
  - \def\bbeta{\bm{\beta}}
  - \def\bSigma{\bm{\Sigma}}
  - \def\Var{\text{Var}}
  - \def\var{\text{Var}}
  - \def\bOmega{\bm{\Omega}}
  - \def\bLambda{\bm{\Lambda}}
  - \let\mc\multicolumn
  - \def\hl{\color[RGB]{6,84,182}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
source("nicefigs.R")
options(digits=4, width=55)
```

# Hierarchical and grouped time series

## Forecasting the PBS
\full{pills}

## ATC drug classification
\fontsize{13}{13}\sf

\hspace*{-0.5cm}\begin{tabular}{lp{10.7cm}}
A &Alimentary tract and metabolism\\
B &Blood and blood forming organs\\
C &Cardiovascular system\\
D &Dermatologicals\\
G &Genito-urinary system and sex hormones\\
H &Systemic hormonal preparations, excluding sex hormones and insulins\\
J &Anti-infectives for systemic use\\
L &Antineoplastic and immunomodulating agents\\
M &Musculo-skeletal system\\
N &Nervous system\\
P &Antiparasitic products, insecticides and repellents\\
R &Respiratory system\\
S &Sensory organs\\
V &Various
\end{tabular}

## ATC drug classification
\hspace*{-1cm}\begin{minipage}{9.6cm}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,font=\small,draw,fill=red!15]
\tikzstyle[level font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 1}=[font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[font=\small,set style={{every node}+=[fill=yellow]}]
\node[label=right:\hspace*{-0.cm}Alimentary tract and metabolism,label=left:\textbf{14 classes}\hspace*{0.3cm}]{A}[edge from parent fork down]
   child {node[label=right:\hspace*{-0.0cm}Drugs used in diabetes,label=left:\textbf{84 classes}]{A10}
     child {node[label=right:\hspace*{-0.35cm}Blood glucose lowering drugs]{A10B}
       child {node[label=right:\hspace*{0.1cm}Biguanides]{A10BA}
         child {node[label=right:\hspace*{-0.15cm}Metformin]{A10BA02}
 }}}};
 \end{tikzpicture}
\end{minipage}

## Australian tourism
\full{regions1_with_labels}
\only<2>{\begin{textblock}{10}(1,1.5)\small
\begin{block}{}
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor night from 1998:Q1 -- 2013:Q4
    \item From: \textit{National Visitor Survey}, based on annual interviews of 120,000 Australians aged 15+, collected by Tourism Research Australia.
    \item Split by 7 states, 27 zones and 76 regions (a geographical hierarchy)
    \item Also split by purpose of travel
      \begin{itemize}
        \item Holiday
        \item Visiting friends and relatives (VFR)
        \item Business
        \item Other
      \end{itemize}
    \item 304 bottom-level series
  \end{itemize}
\end{block}
\end{textblock}}

## Spectacle sales

\placefig{1}{1.4}{width=9.5cm}{spectacles.jpg}
\vspace*{4.8cm}

 * Monthly UK sales data from 2000 -- 2014
 * Provided by a large spectacle manufacturer
 * Split by brand (26), gender (3), price range (6), materials (4), and stores (600)
 * About 1 million bottom-level series

## Hierarchical time series
\fontsize{13}{14}\sf

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}\vspace*{-0.2cm}

 * Pharmaceutical sales
 * Tourism demand by state and region

## Grouped time series
\fontsize{13}{14}\sf

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}

 * Spectacle sales by brand, gender, stores, etc.
 * Tourism by state and purpose of travel

## The problem
\fontsize{13}{14}\sf

\begin{alertblock}{}
\begin{enumerate}\tightlist
 \item How to forecast time series at all nodes such that the forecasts add up in the same way as the original data?
 \item Can we exploit relationships between the series to improve the forecasts?
\end{enumerate}
\end{alertblock}\pause

### The solution

1. Forecast all series at all levels of aggregation using an automatic forecasting algorithm.\newline (e.g., `ets`, `auto.arima`, ...)
2. Reconcile the resulting forecasts so they add up correctly using least squares optimization (i.e., find closest reconciled forecasts to the original forecasts).
3. This is available in the \textbf{hts} package in R.

# hts package for R

## hts package for R
\placefig{9.5}{1}{width=2cm}{Rlogo}

\begin{textblock}{12.}(0.5,1.5)\fontsize{11}{14}\sf
\alert{hts: Hierarchical and Grouped Time Series}

Methods for analysing and forecasting hierarchical and grouped time series\vspace*{0.1cm}

\tabcolsep=0.1cm
\begin{tabular}{@{}ll}
Version: & 5.1.5\\
Depends: & R ($\ge$ 3.2.0), forecast ($\ge$ 8.1) \\
Imports: & SparseM, Matrix, matrixcalc, parallel, utils, methods, graphics, grDevices, stats\\
LinkingTo: & Rcpp ($\ge$ 0.11.0), RcppEigen\\
Suggests: &testthat, knitr, rmarkdown\\
Published: & 2018-03-26\\
Author: &Rob J Hyndman, Alan Lee, Earo Wang, Shanika Wickramasuriya\\
Maintainer: &Rob J Hyndman $<$Earo.Wang at gmail.com$>$\\
BugReports: & \url{https://github.com/earowang/hts/issues} \\
License: & GPL ($\ge2$) \\
URL: & \url{http://pkg.earo.me/hts}
\end{tabular}
\end{textblock}

## Example using R
\fontsize{10}{10}\sf

```r
library(hts)

# bts is a matrix containing the bottom level time series
# nodes describes the hierarchical structure
y <- hts(bts, nodes=list(2, c(3,2)))
```

\only<2>{\begin{textblock}{5}(7,3.5){\begin{block}{}\centering\small
\begin{tikzpicture}[level distance=1cm]
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.01cm]
\tikzstyle[sibling distance=4cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\scriptsize,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 1}=[sibling distance=30mm,font=\footnotesize,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
   child {node {AZ}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\end{block}}\end{textblock}}

\vspace*{10cm}

## Example using R
\fontsize{10}{10}\sf

```r
library(hts)

# bts is a matrix containing the bottom level time series
# nodes describes the hierarchical structure
y <- hts(bts, nodes=list(2, c(3,2)))
```

\begin{textblock}{5}(7,3.5){\begin{block}{}\centering\small
\begin{tikzpicture}[level distance=1cm]
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.01cm]
\tikzstyle[sibling distance=4cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\scriptsize,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 1}=[sibling distance=30mm,font=\footnotesize,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
   child {node {AZ}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\end{block}}\end{textblock}

\vspace*{3.4cm}

```r
# Forecast 10-step-ahead using WLS combination method
# ETS used for each series by default
fc <- forecast(y, h=10)
```

\vspace*{10cm}

## forecast.gts() function
\fontsize{10}{10}\sf

\alert{Usage}

```r
forecast(object, h,
  method = c("comb", "bu", "mo", "tdgsf", "tdgsa", "tdfp"),
  fmethod = c("ets", "rw", "arima"),
  weights = c("wls", "ols", "mint", "nseries"),
  covariance = c("shr","sam"),
  positive = TRUE,
  parallel = FALSE, num.cores = 2, ...)
```

\vspace*{-0.2cm}

\alert{Arguments}\vspace*{-0.2cm}

\begin{tabular}{lp{9.4cm}}
\Verb|object| & Hierarchical time series object of class \texttt{gts}.\\
\Verb|h| & Forecast horizon\\
\Verb|method| & Method for distributing forecasts within the hierarchy. \\
\Verb|fmethod| & Forecasting method to use\\
\Verb|weights| & Weights used for "optimal combination" method. \\
\Verb|covariance| & Shrinkage estimator or sample estimator for GLS covariance.\\
\Verb|positive| & If TRUE, forecasts are forced to be strictly positive\\
\Verb|parallel| & If TRUE, allow parallel processing\\
\Verb|num.cores| & If parallel = TRUE, specify how many cores to be used
\end{tabular}

# Application: Australian tourism

## Australian tourism
\full{regions1_with_labels}
\only<3-4>{\begin{textblock}{6.8}(2,2)
\begin{block}{Hierarchy:}
  \begin{itemize}
      \item  States (7)
      \item  Zones (27)
      \item  Regions (82)
  \end{itemize}
\end{block}
\end{textblock}}
\only<4>{\begin{textblock}{6.8}(2,6)
\begin{block}{Base forecasts}
ETS (exponential smoothing) models
\end{block}\end{textblock}}

\only<2>{\begin{textblock}{10.}(1.4,2)
\begin{block}{Domestic visitor nights}
Quarterly data: 1998 -- 2006.\\
From: \textit{National Visitor Survey}, based on annual interviews of 120,000 Australians aged 15+, collected by Tourism Research Australia.
\end{block}
\end{textblock}}

## Base forecasts
\only<1>{\full{austourism1}}
\only<2>{\full{austourism2}}
\only<3>{\full{austourism3}}
\only<4>{\full{austourism4}}
\only<5>{\full{austourism5}}
\only<6>{\full{austourism6}}
\only<7>{\full{austourism7}}
\only<8>{\full{austourism8}}
\only<9>{\full{austourism9}}

## Reconciled forecasts
\only<1>{\full{Australia}}
\only<2>{\full{States}}
\only<3>{\full{Capitals}}

## Forecast evaluation

  * Select models using all observations;
  * Re-estimate models using first 12 observations and generate 1- to 8-step-ahead forecasts;
  * Increase sample size one observation at a time, re-estimate models, generate forecasts until the end of the sample;
  * In total 24 1-step-ahead, 23 2-steps-ahead, up to 17 8-steps-ahead for forecast evaluation.

## Forecast evaluation

```{r rorigin, include=FALSE}
for(i in 2:6)
{
  fname <- paste("rollingorigin",i,sep="")
  savepdf(fname,height=9.5,width=15)
  plot(0,0,xlim=c(0,28),ylim=c(0,1),
       xaxt="n",yaxt="n",bty="n",xlab="",ylab="",type="n")
  for(j in 1:20)
  {
    test <- (6+j):26
    train <- 1:(5+j)
    arrows(0,1-j/20,27,1-j/20,0.05)
    points(train,rep(1-j/20,length(train)),pch=19,col="blue")
    if(length(test) >= i)
      points(test[i], 1-j/20, pch=19, col="red")
    if(length(test) >= i)
      points(test[-i], rep(1-j/20,length(test)-1), pch=19, col="gray")
    else
      points(test, rep(1-j/20,length(test)), pch=19, col="gray")
  }
  text(28,.95,"time")
  endpdf()
}
for(k in 1:20) {
  fname <- paste("rorigin",k,sep="")
  savepdf(fname,height=9.5,width=15)
  plot(0,0,xlim=c(0,28),ylim=c(0,1),
     xaxt="n",yaxt="n",bty="n",xlab="",ylab="",type="n")
  for(j in 1:k) {
    test <- (6+j):26
    train <- 1:(5+j)
    arrows(0,1-j/20,27,1-j/20,0.05)
    points(train,rep(1-j/20,length(train)),pch=19,col="blue")
    if(length(test) >= 1) {
      points(test[1], 1-j/20, pch=19, col="red")
      points(test[-1], rep(1-j/20,length(test)-1), pch=19, col="gray")
    }
    else {
      points(test, rep(1-j/20,length(test)), pch=19, col="gray")
    }
  }
  text(28,.95,"time")
  endpdf()
}
```

\begin{textblock}{14}(0.3,1.4)
\textbf{\textcolor{blue}{Training sets}} \hspace*{3cm}
\textbf{\textcolor{red}{Test sets
\only<1-20>{$h=1$}%
\only<21>{$h=2$}%
\only<22>{$h=3$}%
\only<23>{$h=4$}%
\only<24>{$h=5$}%
\only<25>{$h=6$}%
}}

\end{textblock}
\only<1>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin1}}
\only<2>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin2}}
\only<3>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin3}}
\only<4>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin4}}
\only<5>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin5}}
\only<6>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin6}}
\only<7>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin7}}
\only<8>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin8}}
\only<9>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin9}}
\only<10>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin10}}
\only<11>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin11}}
\only<12>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin12}}
\only<13>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin13}}
\only<14>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin14}}
\only<15>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin15}}
\only<16>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin16}}
\only<17>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin17}}
\only<18>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin18}}
\only<19>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin19}}
\only<20>{\placefig{0.0}{2.0}{width=12.6cm}{rorigin20}}

\only<21>{\placefig{0.0}{2.0}{width=12.6cm}{rollingorigin2}}
\only<22>{\placefig{0.0}{2.0}{width=12.6cm}{rollingorigin3}}
\only<23>{\placefig{0.0}{2.0}{width=12.6cm}{rollingorigin4}}
\only<24>{\placefig{0.0}{2.0}{width=12.6cm}{rollingorigin5}}
\only<25>{\placefig{0.0}{2.0}{width=12.6cm}{rollingorigin6}}

## Hierarchy: states, zones, regions
\fontsize{10}{10.5}\sf\tabcolsep=0.12cm\vspace*{-0.6cm}
\hspace*{-0.2cm}\begin{tabular}{lrrrrrrr}
            & \multicolumn{6}{c}{\bf Forecast horizon} & \\
\textbf{RMSE} & $h=1$  & $h=2$ & $h=3$ & $h=4$ & $h=5$ & $h=6$ & \textbf{Ave}\\
\midrule
\multicolumn{8}{c}{\bf\alert{Australia}}\\
Base   & 1762.04    & 1770.29    & 1766.02     & 1818.82    & 1705.35    & 1721.17    & \bf 1757.28 \\
Bottom & 1736.92    & 1742.69    & 1722.79     & 1752.74    & 1666.73    & 1687.43    & \bf 1718.22 \\
WLS    & 1705.21    & 1715.87    & \hl 1703.75 & 1729.56    & 1627.79    & \hl1661.24 & \bf 1690.57 \\
GLS    & \hl1704.64 & \hl1715.60 & 1705.31     & \hl1729.04 & \hl1626.36 & 1661.64    & \bf \hl 1690.43 \\[0.1cm]
\multicolumn{8}{c}{\bf\alert{States}}\\
Base   & 399.77     & 404.16     & 401.92      & 407.26     & 395.38     & 401.17     & \bf 401.61 \\
Bottom & 404.29     & 406.95     & 404.96      & 409.02     & 399.80     & 401.55     & \bf 404.43 \\
WLS    & \hl 398.84 & \hl 402.12 & \hl 400.71  & \hl 405.03 & 394.76     & 398.23     & \bf \hl 399.95 \\
GLS    & \hl 398.84 & 402.16     & 400.86      & \hl 405.03 & \hl 394.59 & \hl 398.22 & \bf \hl 399.95 \\[0.1cm]
\multicolumn{8}{c}{\bf\alert{Regions}}\\
Base   & 93.15      & 93.38      & 93.45       & 93.79      & 93.50      & 93.56      & \bf 93.47 \\
Bottom & 93.15      & 93.38      & 93.45       & 93.79      & 93.50      & 93.56      & \bf 93.47 \\
WLS    & 93.02      & 93.32      & 93.38       & 93.72      & 93.39      & 93.53      & \bf 93.39 \\
GLS    & \hl 92.98  & \hl 93.27  & \hl 93.34   & \hl 93.66  & \hl 93.34  & \hl 93.46  & \bf \hl 93.34
\end{tabular}

# Optimal forecast reconciliation

## Hierarchical time series

\begin{minipage}{4cm}\vspace*{0.2cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\only<2->{\begin{textblock}{6.3}(6,1)\small
\begin{itemize}\itemsep=0cm\parskip=0cm
\item[ $ y_{t}: $] observed aggregate of all series at time
$t$.
\item[ $ y_{X,t}: $] observation on series $X$ at time $t$.
\item[ $ \bm{b}_{t}: $] vector of all series at bottom level
in time $t$.
\end{itemize}
\end{textblock}}\vspace*{0.6cm}
\only<3->{
$\bY_{t}= \begin{pmatrix}
  y_{t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix} = \only<3>{\hspace*{0.01cm}\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}\only<4->{{\color{red}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}}\only<3>{\hspace*{0.08cm}}\only<3>{\hspace*{-0.1cm}\begin{pmatrix}Y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}\rule{0cm}{1.6cm}
                \only<4->{\hspace*{0.08cm}{\color{blue}\underbrace{\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}_{\bm{b}_{t}}}}$}

\vspace*{-0.6cm}

\only<4>{\hspace*{8cm}\colorbox[gray]{.8}{$\bY_{t}=\color{red}\bS\color{blue}\bm{b}_{t}$}}

\vspace*{10cm}

## Hierarchical time series

\begin{block}{}\hspace*{.6cm}{\centering\small
\begin{tikzpicture}[level distance=1cm]
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.01cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\scriptsize,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 1}=[sibling distance=40mm,font=\footnotesize,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
   child {node {AZ}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
   child {node {BZ}}
 }
 child {node {C}
   child {node {CX}}
   child {node {CY}}
   child {node {CZ}}
 };
\end{tikzpicture}}
\end{block}\vspace*{0.1cm}\pause

\hbox{$\by_{t}= {\footnotesize\begin{pmatrix}
    y_t\\[-0.15cm]
    y_{A,t}\\[-0.15cm]
    y_{B,t}\\[-0.15cm]
    y_{C,t}\\[-0.15cm]
    y_{AX,t}\\[-0.15cm]
    y_{AY,t}\\[-0.15cm]
    y_{AZ,t}\\[-0.15cm]
    y_{BX,t}\\[-0.15cm]
    y_{BY,t}\\[-0.15cm]
    y_{BZ,t}\\[-0.15cm]
    y_{CX,t}\\[-0.15cm]
    y_{CY,t}\\[-0.15cm]
    y_{CZ,t}\end{pmatrix}}=
    {\color{red}\underbrace{\footnotesize\begin{pmatrix}
                1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\[-0.15cm]
                1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1\\[-0.15cm]
                1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0\\[-0.15cm]
                0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
             \end{pmatrix}}_{\bS}}{\color{blue}\underbrace{\footnotesize\begin{pmatrix}
    y_{AX,t}\\[-0.15cm]
    y_{AY,t}\\[-0.15cm]
    y_{AZ,t}\\[-0.15cm]
    y_{BX,t}\\[-0.15cm]
    y_{BY,t}\\[-0.15cm]
    y_{BZ,t}\\[-0.15cm]
    y_{CX,t}\\[-0.15cm]
    y_{CY,t}\\[-0.15cm]
    y_{CZ,t}\end{pmatrix}}_{\bm{b}_{t}}}$}

\vspace*{10cm}

\only<3>{\begin{textblock}{3}(10.5,8)\colorbox[gray]{.8}{$\by_{t}=\color{red}\bS\color{blue}\bm{b}_{t}$}\end{textblock}}

## Grouped time series

\begin{block}{}
\begin{center}\small
\tikzstyle{every node}=[inner sep=2pt]
\begin{tikzpicture}
    \matrix[ampersand replacement=\&,column sep=0.3cm] {
        \node[ellipse,draw,fill=yellow,font=\scriptsize,distance=1cm] {AX};~ \&
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {AY};~ \&
        \node[ellipse,draw,fill=blue!15] {A}; \\[0.3cm]
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {BX};~ \&
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {BY};~ \&
        \node[ellipse,draw,fill=blue!15] {B}; \\[0.3cm]
        \node[ellipse,draw,fill=blue!15] {X};~ \&
        \node[ellipse,draw,fill=blue!15] {Y};~ \&
        \node[ellipse,draw,fill=red!15] {Total}; \\
};
\end{tikzpicture}
\end{center}
\end{block}\vspace*{0.0cm}\pause\fontsize{13}{14}\sf

\hbox{$\by_{t}= {\small\begin{pmatrix}
    y_t\\[-0.15cm]
    y_{A,t}\\[-0.15cm]
    y_{B,t}\\[-0.15cm]
    y_{X,t}\\[-0.15cm]
    y_{Y,t}\\[-0.15cm]
    y_{AX,t}\\[-0.15cm]
    y_{AY,t}\\[-0.15cm]
    y_{BX,t}\\[-0.15cm]
    y_{BY,t}
    \end{pmatrix}}=
    \color{red}\underbrace{\small\begin{pmatrix}
                1 & 1 & 1 & 1 \\[-0.15cm]
                1 & 1 & 0 & 0 \\[-0.15cm]
                0 & 0 & 1 & 1 \\[-0.15cm]
                1 & 0 & 1 & 0 \\[-0.15cm]
                0 & 1 & 0 & 1 \\[-0.15cm]
                1 & 0 & 0 & 0 \\[-0.15cm]
                0 & 1 & 0 & 0 \\[-0.15cm]
                0 & 0 & 1 & 0 \\[-0.15cm]
                0 & 0 & 0 & 1
             \end{pmatrix}}_{\bS} \color{blue}\underbrace{\small\begin{pmatrix}
    y_{AX,t}\\[-0.15cm]
    y_{AY,t}\\[-0.15cm]
    y_{BX,t}\\[-0.15cm]
    y_{BY,t}
    \end{pmatrix}}_{\bm{b}_{t}}$}

\vspace*{-1cm}

\only<3>{\begin{textblock}{3}(10.5,8)\colorbox[gray]{.8}{$\by_{t}=\color{red}\bS\color{blue}\bm{b}_{t}$}\end{textblock}}

\vspace*{10cm}

## Hierarchical and grouped time series

Every collection of time series with aggregation constraints can be written as
\begin{block}{}
\centerline{$\by_{t}=\bS\bm{b}_{t}$}
\end{block}
where

 * $\by_t$ is a vector of all series at time $t$
 * $\bm{b}_t$ is a vector of the most disaggregated series at time $t$
 * $\bS$ is a ``summing matrix'' containing the aggregation constraints.

## Forecasting notation

Let $\hat{\by}_n(h)$ be vector of initial $h$-step forecasts, made at time $n$, stacked in same order as $\by_t$. \pause\newline  (In general, they will not ``add up''.)\pause

\begin{block}{}
Reconciled forecasts must be of the form:
$$\tilde{\by}_{n}(h)=\bS\bm{P}\hat{\by}_{n}(h)$$
for some matrix $\bm{P}$.
\end{block}\pause

 * $\bm{P}$ extracts and combines base forecasts $\hat{\by}_{n}(h)$ to get bottom-level forecasts.
 * $\bS$ adds them up

## Optimal combination forecasts
\fontsize{14}{15}\sf

\begin{alertblock}{Main result}
The best (minimum sum of variances) unbiased forecasts are obtained when
$\bm{P} = (\bS'\bSigma^{-1}_{h}\bS)^{-1}\bS'\bSigma^{-1}_{h}$,
where $\bSigma_h$ is the $h$-step base forecast error covariance matrix.
\end{alertblock}

\pause

\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{n}(h)}
=\bS(\bS'\bSigma^{-1}_{h}\bS)^{-1}\bS'\bSigma^{-1}_{h}\textcolor{blue}{\hat{\by}_{n}(h)}$}
\end{block}\fontsize{14}{15}\sf

\alert{\textbf{Problem:}} $\bSigma_h$ hard to estimate, especially for $h>1$.

\alert{Solutions:}\vspace*{-0.4cm}

 * Ignore $\bSigma_h$ (OLS)
 * Assume $\bSigma_h$ diagonal (WLS) [Default in `hts`]
 * Try to estimate $\bSigma_h$ (GLS)

## Features
\fontsize{15}{17}\sf

 * Covariates can be included in initial forecasts.
 * Adjustments can be made to initial forecasts at any level.
 * Very simple and flexible method. Can work with *any* hierarchical or grouped time series.
 * Conceptually easy to implement: regression of base forecasts on structure matrix.

# Lab Session 22
##
\fontsize{48}{60}\sf\centering
**Lab Session 22**

# Temporal hierarchies

## Temporal hierarchies

\begin{center}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=60mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=30mm,font=\small,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 3}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=green]}]
\node{Annual}[edge from parent fork down]
 child {node {Semi-Annual$_1$}
   child {node {Q$_1$}
   }
   child {node {Q$_2$}
   }
 }
 child {node {Semi-Annual$_2$}
   child {node {Q$_3$}
   }
   child {node {Q$_4$}
   }
 };
\end{tikzpicture}
\end{center}
\pause
\begin{alertblock}{Basic idea:}
\begin{itemize}
\item Forecast series at each available frequency.
\item Optimally reconcile forecasts within the same year.
\end{itemize}
\end{alertblock}

## Monthly series

\only<1>{
\hspace*{-0.5cm}\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=60mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=30mm,font=\small,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 3}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=green]}]
\node{Annual}[edge from parent fork down]
 child {node {Semi-Annual$_1$}
   child {node {Q$_1$}
     child {node {M$_1$}}
     child {node {M$_2$}}
     child {node {M$_3$}}
   }
   child {node {Q$_2$}
     child {node {M$_4$}}
     child {node {M$_5$}}
     child {node {M$_6$}}
   }
 }
 child {node {Semi-Annual$_2$}
   child {node {Q$_3$}
     child {node {M$_7$}}
     child {node {M$_8$}}
     child {node {M$_9$}}
   }
   child {node {Q$_4$}
     child {node {M$_{10}$}}
     child {node {M$_{11}$}}
     child {node {M$_{12}$}}
   }
 };
\end{tikzpicture}}

\only<2->{
\hspace*{-0.5cm}\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=40mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=20mm,font=\small,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 3}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=green]}]
\node{Annual}[edge from parent fork down]
 child {node {FourM$_1$}
   child {node {BiM$_1$}
     child {node {M$_1$}}
     child {node {M$_2$}}
   }
   child {node {BiM$_2$}
     child {node {M$_3$}}
     child {node {M$_4$}}
   }
 }
 child {node {FourM$_2$}
   child {node {BiM$_3$}
     child {node {M$_5$}}
     child {node {M$_6$}}
   }
   child {node {BiM$_4$}
     child {node {M$_7$}}
     child {node {M$_8$}}
   }
 }
  child {node {FourM$_3$}
   child {node {BiM$_5$}
     child {node {M$_9$}}
     child {node {M$_{10}$}}
   }
   child {node {BiM$_6$}
     child {node {M$_{11}$}}
     child {node {M$_{12}$}}
   }
 };
\end{tikzpicture}}

\biz\tightlist
\item $k=2,4,12$ nodes
\item $k=3,6,12$ nodes
\item Why not $k=2,3,4,6,12$ nodes?
\eiz

## Monthly data
\fontsize{11}{11}\sf\vspace*{-0.4cm}

$${\underbrace{
\begin{pmatrix}
    A\\
    SemiA_{1}\\
    SemiA_{2}\\
    FourM_{1}\\
    FourM_{2}\\
    FourM_{3}\\
    Q_{1}\\
    \vdots\\
    Q_{4}\\
    BiM_{1}\\
    \vdots\\
    BiM_{6}\\
    M_{1}\\
    \vdots\\
    M_{12}
    \end{pmatrix}}_{(28\times1)}}=
    {\color{red}\underbrace{
    \begin{pmatrix}
                1 & 1 & 1 & 1 & 1~~~1~~~1~~~1 & 1 & 1& 1& 1\\
                1 & 1 & 1 & 1 & 1~~~1~~~0~~~0 & 0 & 0& 0& 0\\
                0 & 0 & 0 & 0 & 0~~~0~~~1~~~1 & 1 & 1& 1& 1\\
                1 & 1 & 1 & 1 & 0~~~0~~~0~~~0 & 0 & 0& 0& 0\\
                0 & 0 & 0 & 0 & 1~~~1~~~1~~~1 & 0 & 0& 0& 0\\
                0 & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 1 & 1& 1& 1\\
                1 & 1 & 1 & 0 & 0~~~0~~~0~~~0 & 0 & 0& 0& 0\\
                  &   &   &   & \vdots        &   &  &  &  \\
                0 & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 1& 1& 1\\
                1 & 1 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0& 0& 0\\
                  &   &   &   & \vdots        &   &  &  &  \\
                0 & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0& 1& 1\\
                  &   &   &  &  &  &  &  & \\
                \phantom{\vdots}  &   &   &  & \bm{I}_{12} &  &  &  & \\
                  &   &   &  &  &  &  &  & \\
                \end{pmatrix}}_{\bS}}{\color{blue}\underbrace{\begin{pmatrix}
    M_{1}\\
    M_{2}\\
    M_{3}\\
    M_{4}\\
    M_{5}\\
    M_{6}\\
    M_{7}\\
    M_{8}\\
    M_{9}\\
    M_{10}\\
    M_{11}\\
    M_{12}\end{pmatrix}}_{\bm{b}_{t}}}$$

\vspace*{10cm}

## In general
\vspace{-.3cm}

For a time series  $y_1,\dots,y_T$, observed at frequency $m$, we generate aggregate series
\begin{alertblock}{}%\vspace*{-0.4cm}
\[
y_j^{\left[k\right]} = \sum^{jk}_{t=1+(j-1)k}{y_t},\qquad \text{for $j = 1,\dots,\lfloor T/k\rfloor$}
\]
\end{alertblock}

 * $k \in F(m)=\{\text{factors of $m$}\}$.
 * A single unique hierarchy is only possible when there are no coprime pairs in $F(m)$.
 * $M_k=m/k$ is seasonal period of aggregated series.

## UK Accidents and Emergency Demand}
\full{AEexample}

\begin{textblock}{12}(0.3,8.5)
\textcolor{red}{-- -- -- -- base} \hspace*{1cm}
\textcolor{blue}{\raisebox{0.5ex}{\rule{1.5cm}{1pt}} reconciled}
\end{textblock}

## UK Accidents and Emergency Demand
\fontsize{13}{15}\sf

1. Type 1 Departments --- Major A\&E
1. Type 2 Departments --- Single Specialty
1. Type 3 Departments --- Other A\&E/Minor Injury
1. Total Attendances
1. Type 1 Departments --- Major A\&E $>4$ hrs
1. Type 2 Departments --- Single Specialty $>4$ hrs
1. Type 3 Departments --- Other A\&E/Minor Injury $>4$ hrs
1. Total Attendances $>4$  hrs
1. Emergency Admissions via Type 1 A\&E
1. Total Emergency Admissions via A\&E
1. Other Emergency Admissions (i.e., not via A\&E)
1. Total Emergency Admissions
1. Number of patients spending $>4$ hrs from decision to admission

## UK Accidents and Emergency Demand
\vspace*{-0.3cm}

 * **Minimum training set**: all data except the last year
 * Base forecasts using \texttt{auto.arima()}.
 * Mean Absolute Scaled Errors for 1, 4 and 13 weeks ahead using a rolling origin.

\pause\fontsize{12}{14}\sf
###
\centering
\begin{tabular}{lrccr}
\bf Aggr. Level     & $h$   & \bf Base  & \bf Reconciled & \bf Change \\
\midrule
Weekly & 1      & 1.6   & 1.3   & $-17.2\%$ \\
Weekly & 4      & 1.9   & 1.5   & $-18.6\%$ \\
Weekly & 13     & 2.3   & 1.9   & $-16.2\%$ \\
Weekly & 1--52  & 2.0   & 1.9   & $-5.0\%$ \\
Annual & 1      & 3.4   & 1.9   & $-42.9\%$
\end{tabular}


## thief package for R

\alert{\fontsize{20}{25}\sffamily \alert{thief:} \alert{T}emporal \alert{HIE}rarchical \alert{F}orecasting}\vspace*{.6cm}\pause

\begin{block}{Install from CRAN}
  \texttt{install.packages("thief")}
\end{block}

\begin{alertblock}{Usage}
  \texttt{library(thief)}\\
  \texttt{thief(y)}
\end{alertblock}

# Lab session 23

##
\fontsize{48}{60}\sf\centering
**Lab Session 23**
