---
title: "Forecasting: principles and practice"
author: "Rob J Hyndman"
date: "3.4&nbsp; Extras"
fontsize: 14pt
output:
  beamer_presentation:
    fig_width: 7
    fig_height: 4.3
    highlight: tango
    theme: metropolis
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache=TRUE,
  warning=FALSE,
  message=FALSE)
library(fpp2)
options(digits=4, width=55)
```


# Bagged ETS

## Bagged ETS

\fontsize{10}{10}\sf

### Algorithm: Generating bootstrapped series
```r
bootstrap ← function(ts, num.boot) {
  lambda ← BoxCox.lambda(ts, min=0, max=1)
  ts.bc ← BoxCox(ts, lambda)
  if(ts is seasonal) {
    [trend, seasonal, remainder] ← stl(ts.bc)
  }
  else {
    seasonal ← 0
    [trend, remainder] ← loess(ts.bc)
  }
  recon.series[1] ← ts
  for(i in 2:num.boot) {
    boot.sample[i] ← MBB(remainder)
    recon.series.bc[i] ← trend + seasonal + boot.sample[i]
    recon.series[i] ← InvBoxCox(recon.series.bc[i], lambda)
  }
  return(recon.series)
```


## Bagged ETS

```{r baggedets1, echo=FALSE}
library(Mcomp)
df <- cbind(
  'M495' = M3[[1896]]$x,
  'Box-Cox transformed' = BoxCox(M3[[1896]]$x, BoxCox.lambda(M3[[1896]]$x))
)
autoplot(df, facets=TRUE)
```

## Bagged ETS

```{r  baggedets2, echo=FALSE}
df[,2] %>% stl(s.window='periodic') %>% autoplot
```

## Bagged ETS

```{r baggedets3, echo=FALSE}
fit <- bld.mbb.bootstrap(Mcomp::M3[[1896]]$x, 11)
lambda <- BoxCox.lambda(Mcomp::M3[[1896]]$x)
as.data.frame(fit) %>% as.ts -> z
tsp(z) <- tsp(Mcomp::M3[[1896]]$x)
colnames(z) <- c("Data",paste("B",1:10,sep=""))
```

```{r baggedets4, echo=FALSE}
stlb <- stl(df[,2], s.window='periodic')
stlb$time.series[,3] <- BoxCox(z[,2],lambda) - stlb$time.series[,1] - stlb$time.series[,2]
autoplot(stlb)
```

## Bagged ETS

```{r baggedets5, echo=FALSE}
autoplot(Mcomp::M3[[1896]]$x) +
  xlab("Year") + ylab("M495") +
  ylim(min(z),max(z))
```



## Bagged ETS

```{r baggedets6, echo=FALSE}
autoplot(Mcomp::M3[[1896]]$x) +
  forecast::autolayer(z[,2]) +
  xlab("Year") + ylab("M495") +
  ylim(min(z),max(z)) +
  theme(legend.position="none")
```

## Bagged ETS

```{r baggedets7, echo=FALSE}
autoplot(Mcomp::M3[[1896]]$x) +
  forecast::autolayer(z[,2:11]) +
  xlab("Year") + ylab("M495") +
  ylim(min(z),max(z)) +
  theme(legend.position="none")
```


## Bagged ETS

\fontsize{11}{11}\sf

```{r baggedets8, echo=TRUE, cache=TRUE, fig.height=4}
baggedETS(Mcomp::M3[[1896]]$x) %>%
  forecast %>% autoplot +
  xlab("Year") + ylab("M495")
```


## Bagged ETS

```{r baggedets9, echo=FALSE, cache=TRUE, fig.height=4}
baggedETS(Mcomp::M3[[1896]]$x) %>%
  forecast %>% autoplot +
  xlab("Year") + ylab("M495")
```

 * Intervals show range of point forecasts
 * They are not prediction intervals


# Models for different frequencies

## Models for different frequencies

### Models for annual data

 * ETS, ARIMA, Dynamic regression
\pause

### Models for quarterly data

 * ETS, ARIMA/SARIMA, Dynamic regression, Dynamic harmonic regression, STL+ETS, STL+ARIMA
\pause

### Models for monthly data

 * ETS, ARIMA/SARIMA, Dynamic regression, Dynamic harmonic regression, STL+ETS, STL+ARIMA

## Models for different frequencies

### Models for weekly data

 * ARIMA/SARIMA, Dynamic regression, Dynamic harmonic regression, STL+ETS, STL+ARIMA, TBATS
\pause

### Models for daily, hourly and other sub-daily data

 * ARIMA/SARIMA, Dynamic regression, Dynamic harmonic regression, STL+ETS, STL+ARIMA, TBATS


# Ensuring forecasts stay within limits

## Positive forecasts
\fontsize{12}{12}\sf

```{r, fig.height=3.8}
eggs %>%
  ets(model="AAN", damped=FALSE, lambda=0) %>%
  forecast(h=50, biasadj=TRUE) %>%
  autoplot()
```

## Forecasts constrained to an interval

Suppose egg prices constrained to lie within $a=50$ and $b=400$.

Transform data using scaled logit transform:
$$
y = \log\left(\frac{x-a}{b-x}\right),
$$
where $x$ is on the original scale and $y$ is the transformed data. To reverse the transformation, we will use
$$
x  = \frac{(b-a)e^y}{1+e^y} + a.
$$

## Forecasts constrained to an interval
\fontsize{12}{12}\sf

```r
    # Bounds
    a <- 50
    b <- 400
    # Transform data and fit model
    fit <- log((eggs-a)/(b-eggs)) %>%
      ets(model="AAN", damped=FALSE)
    fc <- forecast(fit, h=50)
    # Back-transform forecasts
    fc[["mean"]] <- (b-a)*exp(fc[["mean"]]) /
      (1+exp(fc[["mean"]])) + a
    fc[["lower"]] <- (b-a)*exp(fc[["lower"]]) /
     (1+exp(fc[["lower"]])) + a
    fc[["upper"]] <- (b-a)*exp(fc[["upper"]]) /
     (1+exp(fc[["upper"]])) + a
    fc[["x"]] <- eggs
    autoplot(fc)
```

## Forecasts constrained to an interval

```{r constrained, echo=FALSE}
    # Bounds
    a <- 50
    b <- 400
    # Transform data and fit model
    fit <- log((eggs-a)/(b-eggs)) %>%
      ets(model="AAN", damped=FALSE)
    fc <- forecast(fit, h=50)
    # Back-transform forecasts
    fc[["mean"]] <- (b-a)*exp(fc[["mean"]]) /
      (1+exp(fc[["mean"]])) + a
    fc[["lower"]] <- (b-a)*exp(fc[["lower"]]) /
     (1+exp(fc[["lower"]])) + a
    fc[["upper"]] <- (b-a)*exp(fc[["upper"]]) /
     (1+exp(fc[["upper"]])) + a
    fc[["x"]] <- eggs
    # Plot result on original scale
    autoplot(fc)
```

# Forecast combinations

## Forecast combinations

### Clemen (1989)
"The results have been virtually unanimous: combining multiple forecasts leads to increased forecast accuracy. \dots In many cases one can make dramatic performance improvements by simply averaging the forecasts."

## Forecast combinations
\fontsize{10}{10}\sf\vspace*{-0.2cm}
```r
train <- window(auscafe, end=c(2012,9))
h <- length(auscafe) - length(train)
ETS <- forecast(ets(train), h=h)
ARIMA <- forecast(auto.arima(train, lambda=0, biasadj=TRUE),
  h=h)
STL <- stlf(train, lambda=0, h=h, biasadj=TRUE)
NNAR <- forecast(nnetar(train), h=h)
TBATS <- forecast(tbats(train, biasadj=TRUE), h=h)
Combination <- (ETS[["mean"]] + ARIMA[["mean"]] +
  STL[["mean"]] + NNAR[["mean"]] + TBATS[["mean"]])/5

autoplot(auscafe) +
  autolayer(ETS, series="ETS", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(STL, series="STL", PI=FALSE) +
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(TBATS, series="TBATS", PI=FALSE) +
  autolayer(Combination, series="Combination") +
  xlab("Year") + ylab("$ billion") +
  ggtitle("Australian monthly expenditure on eating out")
```

## Forecast combinations
\fontsize{10}{10}\sf\vspace*{-0.2cm}
```{r combine1, message=FALSE, warning=FALSE, echo=FALSE}
train <- window(auscafe, end=c(2012,9))
h <- length(auscafe) - length(train)
ETS <- forecast(ets(train), h=h)
ARIMA <- forecast(auto.arima(train, lambda=0, biasadj=TRUE),
  h=h)
STL <- stlf(train, lambda=0, h=h, biasadj=TRUE)
NNAR <- forecast(nnetar(train), h=h)
TBATS <- forecast(tbats(train, biasadj=TRUE), h=h)
Combination <- (ETS[["mean"]] + ARIMA[["mean"]] +
  STL[["mean"]] + NNAR[["mean"]] + TBATS[["mean"]])/5
```

```{r combineplot, dependson="combine1", echo=FALSE, fig.height=4.8}
autoplot(auscafe) +
  autolayer(ETS, series="ETS", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(STL, series="STL", PI=FALSE) +
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(TBATS, series="TBATS", PI=FALSE) +
  autolayer(Combination, series="Combination") +
  xlab("Year") + ylab("$ billion") +
  ggtitle("Australian monthly expenditure on eating out")
```

## Forecast combinations
\fontsize{11}{15}\sf
```{r combineaccuracy, dependson="combine1"}
c(ETS = accuracy(ETS, auscafe)["Test set","RMSE"],
  ARIMA = accuracy(ARIMA, auscafe)["Test set","RMSE"],
  `STL-ETS` = accuracy(STL, auscafe)["Test set","RMSE"],
  NNAR = accuracy(NNAR, auscafe)["Test set","RMSE"],
  TBATS = accuracy(TBATS, auscafe)["Test set","RMSE"],
  Combination =
    accuracy(Combination, auscafe)["Test set","RMSE"])
```

# Prediction intervals for aggregates
## Prediction intervals for aggregates
\fontsize{11}{11}\sf

```{r aggregates}
# First fit a model to the data
fit <- ets(gas/1000)
# Forecast six months ahead
fc <- forecast(fit, h=6)
sum(fc[["mean"]][1:6])
# Simulate 10000 future sample paths
nsim <- 10000
h <- 6
sim <- numeric(nsim)
for(i in seq_len(nsim))
  sim[i] <- sum(simulate(fit, future=TRUE, nsim=h))
mean(sim)
```

## Prediction intervals for aggregates

```{r aggregates3, dependson="aggregates"}
#80% interval:
quantile(sim, prob=c(0.1, 0.9))
#95% interval:
quantile(sim, prob=c(0.025, 0.975))
```


# Backcasting
## Backcasting
\fontsize{11}{11}\sf

```{r backcasting_functions}
# Function to reverse time
reverse_ts <- function(y)
{
  ts(rev(y), start=tsp(y)[1L], frequency=frequency(y))
}
# Function to reverse a forecast
reverse_forecast <- function(object)
{
  h <- length(object[["mean"]])
  f <- frequency(object[["mean"]])
  object[["x"]] <- reverse_ts(object[["x"]])
  object[["mean"]] <- ts(rev(object[["mean"]]),
    end=tsp(object[["x"]])[1L]-1/f, frequency=f)
  object[["lower"]] <- object[["lower"]][h:1L,]
  object[["upper"]] <- object[["upper"]][h:1L,]
  return(object)
}
```


## Backcasting

\fontsize{11}{11}\sf

```r
euretail %>% reverse_ts() %>%
  auto.arima() %>%
  forecast() %>% reverse_forecast() -> bc
autoplot(bc) +
  ggtitle(paste("Backcasts from",bc[["method"]]))
```


```{r backcasting, dependson="backcasting_functions", echo=FALSE, fig.height=3.5}
# Backcast example
euretail %>%
  reverse_ts() %>%
  auto.arima() %>%
  forecast() %>%
  reverse_forecast() -> bc
autoplot(bc) +
  ggtitle(paste("Backcasts from",bc[["method"]]))
```

# Missing values

## Missing values
\fontsize{13}{14}\sf

**Functions which can handle missing values**

 * `auto.arima()`, `Arima()`
 * `tslm()`
 * `nnetar()`

**Models which cannot handle missing values**

 * `ets()`
 * `stl()`
 * `stlf()`
 * `tbats()`

\pause

### What to do?
 1. Model section of data after last missing value.
 2. Estimate missing values with `na.interp()`.

## Missing values
\fontsize{12}{12}\sf
```{r}
autoplot(gold)
```

## Missing values
\fontsize{12}{12}\sf
```{r, fig.height=3}
gold %>% na.interp() %>%
  autoplot(series="Interpolated") +
    autolayer(gold, series="Original") +
    scale_color_manual(
      values=c(`Interpolated`="red",`Original`="gray"))
```

# Outliers

## Outliers

```{r, fig.height=3.4}
autoplot(gold)
```

## Outliers

```{r, fig.height=3.4}
tsoutliers(gold)
```

## Outliers

```{r, fig.height=3.4}
gold %>% tsclean() %>% autoplot()
```