---
title: "High dimensional time series analysis"
date: "robjhyndman.com/hdtsa"
author: "3. Time series features"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: ../header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
library(broom)
```

# STL Features

## Strength of seasonality and trend

\begin{alertblock}{STL decomposition}
\centerline{$y_t = T_t+S_t+R_t$}
\end{alertblock}

### Seasonal strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\right)$$

### Trend strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r features}
tourism %>% features(Trips, feat_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r features-plot, fig.height=4.1}
tourism %>% features(Trips, feat_stl) %>%
  ggplot(aes(x=trend_strength, y=seasonal_strength_year, col=Purpose)) +
    geom_point() + facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{4.7}(7.8,7.3)
\begin{alertblock}{}\fontsize{10}{10}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism %>%
  features(Trips, feat_stl) %>%
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=2.8}
tourism %>%
  right_join(most_seasonal, by = c("State","Region","Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(vars(State,Region,Purpose))
```


## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most trended time series:

```{r extreme3}
most_trended <- tourism %>%
  features(Trips, feat_stl) %>%
  filter(trend_strength == max(trend_strength))
```

\pause\vspace*{-0.3cm}

```{r extreme4, fig.height=2.8}
tourism %>%
  right_join(most_trended, by = c("State","Region","Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(vars(State,Region,Purpose))
```

# Lab Session 5

## Lab Session 5

 * Use ``GGally::ggpairs()`` to look at the relationships between the STL-based features. You might wish to change `seasonal_peak_year` and `seasonal_trough_year` to factors.
 * Which is the peak quarter for holidays in each state?

# Lag plots and autocorrelation

## Example: Beer production
\fontsize{11}{12}\sf

```{r}
new_production <- aus_production %>%
  filter(year(Quarter) >= 1992)
new_production
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6.5, fig.width=6.5, out.width="8cm"}
new_production %>% gg_lag(Beer)
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6.5, fig.width=6.5, out.width="8cm"}
new_production %>% gg_lag(Beer, geom='point')
```

## Lagged scatterplots

  * Each graph shows $y_t$ plotted against $y_{t-k}$ for
different values of $k$.
  * The autocorrelations are the correlations associated
with these scatterplots.

## Autocorrelation

**Covariance** and **correlation**: measure extent of **linear relationship** between two variables ($y$ and $X$).\pause

**Autocovariance** and **autocorrelation**: measure linear relationship between **lagged values** of a time series $y$.\pause

We measure the relationship between:

  * $y_{t}$ and $y_{t-1}$
  * $y_{t}$ and $y_{t-2}$
  * $y_{t}$ and $y_{t-3}$
  * etc.

## Autocorrelation

We denote the sample autocovariance at lag $k$ by $c_k$ and the sample autocorrelation at lag $k$ by $r_k$.  Then define

\begin{block}{}
\begin{align*}
c_k &= \frac{1}{T}\sum_{t=k+1}^T (y_t-\bar{y})(y_{t-k}-\bar{y}) \\[0.cm]
\text{and}\qquad
r_{k} &= c_k/c_0
\end{align*}
\end{block}\pause\small

  * $r_1$ indicates how successive values of  $y$  relate to each other
  * $r_2$ indicates how  $y$ values two periods apart relate to each other
  * $r_k$ is \textit{almost} the same as the sample correlation between $y_t$ and $y_{t-k}$.

## Autocorrelation


Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r, echo=TRUE}
new_production %>% ACF(Beer, lag_max = 9)
```

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r beeracf, fig.height=2.5}
new_production %>% ACF(Beer, lag_max = 9) %>% autoplot()
```

\vspace*{10cm}

## Autocorrelation

  * $r_{4}$  higher than for the other lags. This is due to **the seasonal pattern in the data**: the peaks tend to be **4 quarters** apart and the troughs tend to be **2 quarters** apart.
  * $r_2$ is more negative than for the other lags because troughs tend to be 2 quarters behind peaks.
  * Together, the autocorrelations at lags 1, 2, \dots, make up the \emph{autocorrelation} or ACF.
  * The plot is known as a **correlogram**

## ACF

```{r, fig.height=4, echo=TRUE}
new_production %>% ACF(Beer) %>% autoplot()
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism %>%
  filter(Purpose=="Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r tourismacf}
holidays %>% ACF(Trips)
```

## Australian holidays
\fontsize{9}{10}\sf

```{r tourismacf2, fig.height=6}
holidays %>% ACF(Trips) %>% autoplot()
```

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r featuresacf}
tourism %>% features(Trips, feat_acf)
```

# Dimension reduction for features

```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width=16, height=10) {
  fname <<- paste("figs/",file,".pdf",sep="")
  pdf(fname, width=width/2.54, height=height/2.54, pointsize=10)
  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.6,1.1,1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs="feasts"))
# Compute PCs
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca2", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
savepdf("pca3", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
endpdf()
# Find outliers
outliers <- pcs %>%
  filter(.fittedPC1 > 12 | (.fittedPC1 > 10 & .fittedPC2 > 0))
savepdf("pca4", 18, 10)
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1) +
  geom_point(data=outliers, aes(x=.fittedPC1, y=.fittedPC2), col="black", shape=1, size=3)
endpdf()
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r tourismfeatures}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs="feasts"))
```

```{r echo=FALSE}
tourism_features
```

\begin{textblock}{2.5}(9.6,1.3)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r pcatable}
pcs <- tourism_features %>% select(-State, -Region, -Purpose) %>%
  prcomp(scale=TRUE) %>% augment(tourism_features)
```

```{r echo=FALSE}
pcs
```

\begin{textblock}{2.5}(9.6,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf


\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca3}}
\only<2>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca4}}
\vspace*{10cm}


## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r outliers2, fig.height=3.5}
outliers %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
    facet_grid(vars(State,Region,Purpose)) +
    ggtitle("Outlying time series in PC space") +
    theme(legend.position = "none")
```

# Lab Session 6

## Lab Session 6

* Use a feature-based approach to look for outlying series in `PBS`.

* What is unusual about these series?

