---
title: "High dimensional time series analysis"
date: "robjhyndman.com/hdtsa"
author: "5. Forecast reconciliation"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: ../header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
```

# Hierarchical and grouped time series

## Australian Pharmaceutical Benefits Scheme

\full{pills}

## PBS sales
\fontsize{9}{10}\sf

```{r setwidth, include=FALSE}
fred <- options(width=71)
```

```{r pbs, dependson='setwidth'}
PBS
```

```{r setwidthback, include=FALSE}
options(width=fred$width)
```

## ATC drug classification
\fontsize{12}{13}\sf

\hspace*{-0.5cm}\begin{tabular}{lp{10.7cm}}
A &Alimentary tract and metabolism\\
B &Blood and blood forming organs\\
C &Cardiovascular system\\
D &Dermatologicals\\
G &Genito-urinary system and sex hormones\\
H &Systemic hormonal preparations, excluding sex hormones and insulins\\
J &Anti-infectives for systemic use\\
L &Antineoplastic and immunomodulating agents\\
M &Musculo-skeletal system\\
N &Nervous system\\
P &Antiparasitic products, insecticides and repellents\\
R &Respiratory system\\
S &Sensory organs\\
V &Various
\end{tabular}

## ATC drug classification
\hspace*{-1.5cm}\begin{minipage}{9.6cm}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,font=\small,draw,fill=red!15]
\tikzstyle[level font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 1}=[font=\small,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[font=\small,set style={{every node}+=[fill=yellow]}]
\node[label=right:\hspace*{-0.cm}Alimentary tract and metabolism,label=left:\textbf{ATC1: 14 classes}\hspace*{0.3cm}]{A}[edge from parent fork down]
   child {node[label=right:\hspace*{-0.0cm}Drugs used in diabetes,label=left:\textbf{ATC2: 84 classes}]{A10}
     child {node[label=right:\hspace*{-0.35cm}Blood glucose lowering drugs]{A10B}
       child {node[label=right:\hspace*{0.1cm}Biguanides]{A10BA}
         child {node[label=right:\hspace*{-0.15cm}Metformin]{A10BA02}
 }}}};
 \end{tikzpicture}
\end{minipage}

## Hierarchical time series
\fontsize{13}{14}\sf

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}\vspace*{-0.2cm}

 * PBS sales by ATC groups
 * Tourism demand by states, zones, regions

## Australian tourism
\full{regions1_with_labels}

## Australian tourism
\fontsize{11}{12}\sf

```{r tourism}
tourism
```

## Australian tourism

\begin{textblock}{10}(1,1.5)\small
\begin{block}{}
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor night from 1998:Q1 -- 2013:Q4
    \item From: \textit{National Visitor Survey}, based on annual interviews of 120,000 Australians aged 15+, collected by Tourism Research Australia.
    \item Split by 7 states, 27 zones and 76 regions (a geographical hierarchy)
    \item Also split by purpose of travel
      \begin{itemize}
        \item Holiday
        \item Visiting friends and relatives (VFR)
        \item Business
        \item Other
      \end{itemize}
    \item 304 bottom-level series
  \end{itemize}
\end{block}
\end{textblock}

## Grouped time series
\fontsize{13}{14}\sf

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}

 * Tourism by state and purpose of travel
 * Retail sales by product groups/sub groups, and by countries/regions

## Creating aggregates
\fontsize{7}{7}\sf

```{r pbs_aggregate}
PBS %>%
  aggregate_key(ATC1/ATC2, Scripts = sum(Scripts)) %>%
  filter(Month == yearmonth("1991 Jul")) %>% print(n=18)
```

## Creating aggregates
\fontsize{8}{8}\sf

```{r tourism_aggregate}
tourism %>%
  aggregate_key(Purpose * (State / Region), Trips = sum(Trips)) %>%
  filter(Quarter == yearquarter("1998 Q1")) %>% print(n=15)
```

## Creating aggregates
\fontsize{13}{15}\sf

 * Similar to `summarise()` but using the key structure
 * A grouped structure is specified using `grp1 * grp2`
 * A nested structure is specified via `parent / child`.
 * Groups and nesting can be mixed:

    ```r
    (country/region/city) * (brand/product)
    ```

 * All possible aggregates are produced.
 * These are useful when forecasting at different levels of aggregation.

# Forecast reconciliation

## The problem
\fontsize{13}{14}\sf

\begin{alertblock}{}
\begin{enumerate}\tightlist
 \item How to forecast time series at all nodes such that the forecasts add up in the same way as the original data?
 \item Can we exploit relationships between the series to improve the forecasts?
\end{enumerate}
\end{alertblock}\pause

### The solution

1. Forecast all series at all levels of aggregation using an automatic forecasting algorithm.\newline (e.g., `ETS`, `ARIMA`, ...)
2. Reconcile the resulting forecasts so they add up correctly using least squares optimization (i.e., find closest reconciled forecasts to the original forecasts).
3. This is available using `reconcile()`.

## Forecast reconciliation
\fontsize{10}{11}\sf

```{r tourismets_reconciled, message=FALSE}
tourism %>%
  aggregate_key(Purpose*(State/Region), Trips=sum(Trips)) %>%
  model(ets = ETS(Trips)) %>%
  reconcile(ets_adjusted = min_trace(ets)) %>%
  forecast(h = 2)
```

## Hierarchical and grouped time series

Every collection of time series with aggregation constraints can be written as
\begin{block}{}
\centerline{$\by_{t}=\bS\bm{b}_{t}$}
\end{block}
where

 * $\by_t$ is a vector of all series at time $t$
 * $\bm{b}_t$ is a vector of the most disaggregated series at time $t$
 * $\bS$ is a ``summing matrix'' containing the aggregation constraints.




## Hierarchical time series

\begin{minipage}{4cm}\vspace*{0.2cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\only<2->{\begin{textblock}{6.3}(6,1)\small
\begin{itemize}\itemsep=0cm\parskip=0cm
\item[$ y_{t}: $] observed aggregate of all series at time
$t$.
\item[$ y_{X,t}: $] observation on series $X$ at time $t$.
\item[$ \bm{b}_{t}: $] vector of all series at bottom level
in time $t$.
\end{itemize}
\end{textblock}}\vspace*{0.6cm}
\only<3->{
$\bY_{t}= \begin{pmatrix}
  y_{t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix} = \only<3>{\hspace*{0.01cm}\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}\only<4->{{\color{blue}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}}\only<3>{\hspace*{0.08cm}}\only<3>{\hspace*{-0.1cm}\begin{pmatrix}Y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}\rule{0cm}{1.6cm}
                \only<4->{\hspace*{0.08cm}{\color{red}\underbrace{\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}_{\bm{b}_{t}}}}$}

\vspace*{-0.6cm}

\only<4>{\hspace*{8cm}\colorbox[RGB]{210,210,210}{$\bY_{t}=\color{blue}\bS\color{red}\bm{b}_{t}$}}

\vspace*{10cm}

## Forecasting notation

Let $\hat{\by}_n(h)$ be vector of initial $h$-step forecasts, made at time $n$, stacked in same order as $\by_t$. \pause\newline  (In general, they will not ``add up''.)\pause

\begin{block}{}
Reconciled forecasts must be of the form:
$$\tilde{\by}_{n}(h)=\bS\bm{G}\hat{\by}_{n}(h)$$
for some matrix $\bm{G}$.
\end{block}\pause

 * $\bm{G}$ extracts and combines base forecasts $\hat{\by}_{n}(h)$ to get bottom-level forecasts.
 * $\bS$ adds them up

## Optimal combination forecasts
\fontsize{13}{14}\sf

\begin{alertblock}{Main result}
The best (minimum sum of variances) unbiased forecasts are obtained when
$\bm{G} = (\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}$,
where $\bW_h$ is the $h$-step base forecast error covariance matrix.
\end{alertblock}

\pause

\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{n}(h)}
=\bS(\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}\textcolor{blue}{\hat{\by}_{n}(h)}$}
\end{block}\fontsize{14}{15}\sf

\alert{\textbf{Problem:}} $\bW_h$ hard to estimate, especially for $h>1$.

\alert{Solutions:}\vspace*{-0.4cm}

 * Ignore $\bW_h$ (OLS) [`min_trace(method='ols')`]
 * Assume $\bW_h = k_h\bW_1$ is diagonal (WLS) [`min_trace(method='wls')`]
 * Assume $\bW_h = k_h\bW_1$ and estimate it (GLS) [`min_trace(method='shrink')` (the default)]

## Features
\fontsize{15}{17}\sf

 * Covariates can be included in initial forecasts.
 * Adjustments can be made to initial forecasts at any level.
 * Very simple and flexible method. Can work with *any* hierarchical or grouped time series.
 * Conceptually easy to implement: regression of base forecasts on structure matrix.

# Example: Australian tourism

## Example: Australian tourism
\fontsize{12}{13}\sf

```{r fctourism}
tourism_agg <- tourism %>%
  aggregate_key(Purpose * (State / Region),
                Trips = sum(Trips))
fc <- tourism_agg %>%
  filter_index(. ~ yearquarter("2015 Q4")) %>%
  model(ets = ETS(Trips)) %>%
  reconcile(ets_adjusted = min_trace(ets)) %>%
  forecast(h = "2 years")
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism2, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & is_aggregated(State)) %>%
  autoplot(tourism_agg, level=95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism3, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & State=="Victoria" &
         is_aggregated(Region)) %>%
  autoplot(tourism_agg, level=95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism4, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region=="Melbourne") %>%
  autoplot(tourism_agg, level=95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism5, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region=="Snowy Mountains") %>%
  autoplot(tourism_agg, level=95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism6, dependson='fctourism'}
fc %>%
  filter(Purpose=="Holiday" & Region=="Barossa") %>%
  autoplot(tourism_agg, level=95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism7, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region=="MacDonnell") %>%
  autoplot(tourism_agg, level=95)
```


## Example: Australian tourism
\fontsize{12}{12.5}\sf

```{r fctourismcomb}
fc <- tourism_agg %>%
  filter_index(. ~ yearquarter("2015 Q4")) %>%
  model(
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  ) %>%
  mutate(
    comb = (ets+arima)/2
  ) %>%
  reconcile(
    ets_adj = min_trace(ets),
    arima_adj = min_trace(arima),
    comb_adj = min_trace(comb)
    ) %>%
  forecast(h = "2 years")
```

## Forecast evaluation
\fontsize{10}{11}\sf

```{r fcaccuracy, dependson='fctourismcomb'}
fc %>% accuracy(tourism_agg)
```

## Forecast evaluation
\fontsize{12}{13}\sf

```{r fcaccuracy2, dependson='fctourismcomb'}
fc %>% accuracy(tourism_agg) %>%
  group_by(.model) %>%
  summarise(MASE = mean(MASE)) %>%
  arrange(MASE)
```

# Lab Session 10

## Lab Session 10

* Prepare aggregations of the PBS data by Concession, Type, and ATC1.
* Use forecast reconciliation with the PBS data, using ETS, ARIMA and SNAIVE models, applied to all but the last 3 years of data.
* Which type of model works best?
* Does the reconciliation improve the forecast accuracy?
* Why doesn't the reconcililation make any difference to the SNAIVE forecasts?
