---
title: "Tidy Time Series & Forecasting in R"
date: "robjhyndman.com/workshop2020"
author: "1. Introduction to tsibbles"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
global_economy <- global_economy %>%
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tourism %>%
  mutate(
    State = recode(State,
                   "Australian Capital Territory" = "ACT",
                   "New South Wales"="NSW",
                   "Northern Territory" = "NT",
                   "Queensland" = "QLD",
                   "South Australia" = "SA",
                   "Tasmania" = "TAS",
                   "Victoria"="VIC",
                   "Western Australia" = "WA"
    )
  )
```

# Time series data and tsibbles

## Tidyverts packages

\begin{textblock}{3.8}(8,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{1}{1.4}{width=4cm}{tsibble.png}
\placefig{5}{1.4}{width=4cm}{tsibbledata.png}
\placefig{3}{4.85}{width=4cm}{feasts.png}
\placefig{7}{4.85}{width=4cm}{fable.png}

## Time series data

  - Four-yearly Olympic winning times
  - Annual Google profits
  - Quarterly Australian beer production
  - Monthly rainfall
  - Weekly retail sales
  - Daily IBM stock prices
  - Hourly electricity demand
  - 5-minute freeway traffic counts
  - Time-stamped stock transaction data

## Class packages
\fontsize{13}{14}\sf

```r
# Data manipulation and plotting functions
library(tidyverse)
# Time series manipulation
library(tsibble)
# Forecasting functions
library(fable)
# Time series graphics and statistics
library(feasts)
# Tidy time series data
library(tsibbledata)
```

\pause

```{r}
# All of the above and more
library(fpp3)
```

## `tsibble` objects

\fontsize{10}{11.2}\sf

```{r, echo = TRUE}
global_economy
```

\only<2->{\begin{textblock}{.75}(2.15,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{1.6}(3.28,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.7}(5.5,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measured variables\phantom{dg}\end{alertblock}
\end{textblock}}

## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4-5>{\begin{textblock}{1.5}(7.95,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\only<5>{\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}}

## `tsibble` objects

* A `tsibble` allows storage and manipulation of multiple time series in R.

* It contains:

  + An index: time information about the observation
  + Measured variable(s): numbers of interest
  + Key variable(s): optional unique identifiers for each series

* It works with tidyverse functions.

## The `tsibble` index

### Example
\fontsize{12}{13}\sf

```{r tstable, cache=TRUE}
mydata <- tsibble(year = 2012:2016,
  y = c(123,39,78,52,110), index = year)
mydata
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r tstablemonth, echo=FALSE}
z <- tibble(Month = paste(2019, month.abb[1:5]), Observation = c(50, 23, 34, 30, 25))
#knitr::kable(z, booktabs=TRUE)
```

```{r tstablemonth2}
z
```

## The `tsibble` index

\begin{block}{}
For observations more frequent than once per year, we need to use a time class function on the index.
\end{block}
\fontsize{12}{13}\sf

```{r month-tsibble}
z %>%
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)
```

## The `tsibble` index

Common time index variables can be created with these functions:

###
```{r tstable2, echo=FALSE}
tribble(
  ~ `Frequency`, ~ Function,
  "Annual", "`start:end`",
  "Quarterly", "`yearquarter()`",
  "Monthly", "`yearmonth()`",
  "Weekly", "`yearweek()`",
  "Daily", "`as_date()`, `ymd()`",
  "Sub-daily", "`as_datetime()`"
) %>%
  knitr::kable(booktabs=TRUE)
```

# Example: Australian prison population

## Australian prison population

\full{Beechworth_prison}

## Create a tsibble from a csv
\fontsize{13}{14}\sf

```{r prison, echo=FALSE, warning=FALSE, message=FALSE}
prison <- readr::read_csv("data/prison_population.csv")
prison %>%
  head(15) %>%
  knitr::kable(booktabs=TRUE)
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison2, eval=FALSE}
prison <- readr::read_csv("prison_population.csv")
```

```{r prison2a, dependson="prison", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison3}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date))
```

```{r prison3a, dependson="prison3", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison4}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date)
```

```{r prison4a, dependson="prison4", echo=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison5}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date) %>%
  as_tsibble(index=Quarter,
    key=c(state, gender, legal, indigenous))
```

```{r prison5a, dependson="prison5", echo=FALSE}
prison
```

# Example: Australian pharmaceutical sales

## Australian Pharmaceutical Benefits Scheme

\full{pills}

## Australian Pharmaceutical Benefits Scheme
\begin{block}{}
The \alert{Pharmaceutical Benefits Scheme} (PBS) is the Australian government drugs subsidy scheme.
\end{block}
\pause\fontsize{13.3}{15}\sf

  * Many drugs bought from pharmacies are subsidised to allow more equitable access to modern drugs.
  * The cost to government is determined by the number and types of drugs purchased. Currently nearly 1\% of GDP.
  * The total cost is budgeted based on forecasts of drug usage.
  * Costs are disaggregated by drug type (ATC1 x`r length(unique(PBS$ATC1))` / ATC2 `r length(unique(PBS$ATC2))`), concession category (x`r length(unique(PBS$Concession))`) and patient type (x`r length(unique(PBS$Type))`), giving $84\times2\times2=`r 84*2*2`$ time series.

## Working with `tsibble` objects {-}
\fontsize{8}{10}\sf

```{r wide, include=FALSE}
options(width=78)
```

```{r pbs1, dependson='wide'}
PBS
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `filter()` function to select rows.

\fontsize{8}{10}\sf

```{r pbs2}
PBS %>%
  filter(ATC2=="A10")
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `select()` function to select columns.

\fontsize{8}{10}\sf

```r
PBS %>%
  filter(ATC2=="A10") %>%
  select(Cost)
```

```
Selecting index: "Month"
Error: The result is not a valid tsibble.
Do you need `as_tibble()` to work with data frame?
```

\vspace*{10cm}

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `select()` function to select columns.

\fontsize{8}{10}\sf

```{r pbs3}
PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `summarise()` function to summarise over keys.

\fontsize{8}{10}\sf

```{r pbs4}
PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost))
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs5}
PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost)) %>%
  mutate(total_cost = total_cost/1e6)
```

## Working with `tsibble` objects {-}
\fontsize{12}{14}\sf

We can use the `mutate()` function to create new variables.

\fontsize{8}{10}\sf

```{r pbs6}
PBS %>%
  filter(ATC2=="A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(total_cost = sum(Cost)) %>%
  mutate(total_cost = total_cost/1e6) -> a10
```

```{r a10, echo=FALSE}
a10
```

```{r narrow, include=FALSE}
options(width=60)
```

# Lab Session 1

## Lab Session 1
\fontsize{14}{15}\sf

 1. Download [`tourism.xlsx`](http://robjhyndman.com/data/tourism.xlsx) from [`http://robjhyndman.com/data/tourism.xlsx`](http://robjhyndman.com/data/tourism.xlsx), and read it into R using `read_excel()` from the `readxl` package.
 2. Create a tsibble which is identical to the `tourism` tsibble from the `tsibble` package.
 3. Find what combination of `Region` and `Purpose` had the maximum number of overnight trips on average.
 4. Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.

# Time plots
## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp, fig.height=3}
maxtemp <- vic_elec %>%
  index_by(Day = date(Time)) %>%
  summarise(Temperature = max(Temperature))
maxtemp %>%
  autoplot(Temperature) +
  xlab("Week") + ylab("Max temperature")
```

## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp2, warning=FALSE, message=FALSE, dependson="maxtemp"}
maxtemp %>%
  ggplot(aes(x = Day, y = Temperature)) +
  geom_point() +
  xlab("Week") + ylab("Max temperature")
```

## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp3, warning=FALSE, message=FALSE, dependson="maxtemp", fig.height=2.5}
maxtemp %>%
  ggplot(aes(x = Day, y = 1)) +
  geom_tile(aes(fill = Temperature)) +
  scale_fill_gradient2(low = "navy", mid = "yellow",
                       high = "red", midpoint=28) +
  ylab("") + scale_y_discrete(expand=c(0,0))
```

## Are line plots best?

\full{TemperatureBlanket}

## Ansett airlines

\full{ansettlogo}

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Class=="Economy") %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Airports=="MEL-SYD") %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Airports=="MEL-SYD") %>%
  autoplot(Passengers)
```

\begin{textblock}{2.2}(10,7.3)
\begin{alertblock}{}
Not the real data! Or is it?
\end{alertblock}
\end{textblock}

# Lab Session 2
## Lab Session 2

- Create time plots of the following time series: `Beer` from `aus_production`, `Lynx` from `pelt`, `Close` from `gafa_stock`
- Use `help()` to find out about the data in each series.
- For the last plot, modify the axis labels and title.

