---
title: "Tidy Time Series & Forecasting in R"
date: "robjhyndman.com/workshop2020"
author: "2. Time series graphics"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
```

# Time plots
## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp, fig.height=3}
maxtemp <- vic_elec %>%
  index_by(Day = date(Time)) %>%
  summarise(Temperature = max(Temperature))
maxtemp %>%
  autoplot(Temperature) +
  xlab("Week") + ylab("Max temperature")
```

## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp2, warning=FALSE, message=FALSE, dependson="maxtemp"}
maxtemp %>%
  ggplot(aes(x = Day, y = Temperature)) +
  geom_point() +
  xlab("Week") + ylab("Max temperature")
```

## Are line plots best?
\fontsize{11}{12}\sf

```{r maxtemp3, warning=FALSE, message=FALSE, dependson="maxtemp", fig.height=2.5}
maxtemp %>%
  ggplot(aes(x = Day, y = 1)) +
  geom_tile(aes(fill = Temperature)) +
  scale_fill_gradient2(low = "navy", mid = "yellow",
                       high = "red", midpoint=28) +
  ylab("") + scale_y_discrete(expand=c(0,0))
```

## Are line plots best?

\full{TemperatureBlanket}


## Ansett airlines

\full{ansettlogo}

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Class=="Economy") %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Airports=="MEL-SYD") %>%
  autoplot(Passengers)
```

## Ansett airlines
\fontsize{11}{12}\sf

```{r, echo=TRUE, fig.height=3.8}
ansett %>%
  filter(Airports=="MEL-SYD") %>%
  autoplot(Passengers)
```

\begin{textblock}{2.2}(10,7.3)
\begin{alertblock}{}
Not the real data! Or is it?
\end{alertblock}
\end{textblock}


# Lab Session 2
## Lab Session 2

- Create time plots of the following time series: `Beer` from `aus_production`, `Lynx` from `pelt`, `Close` from `gafa_stock`
- Use `help()` to find out about the data in each series.
- For the last plot, modify the axis labels and title.

# Seasonal plots

## Seasonal plots

  * Data plotted against the individual "seasons" in which the data were observed.  (In this case a "season" is a month.)
  * Something like a time plot except that the data from each season are overlapped.
  * Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
  * In R: `gg_season()`

## Quarterly Australian Beer Production
\fontsize{13}{15}\sf

```{r, fig.height=3}
beer <- aus_production %>%
  select(Quarter, Beer) %>%
  filter(year(Quarter) >= 1992)
beer %>% autoplot(Beer)
```

## Quarterly Australian Beer Production

```{r}
beer %>% gg_season(Beer, labels="right")
```

## Multiple seasonal periods
\fontsize{10}{11}\sf

```{r}
vic_elec
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r}
vic_elec %>% gg_season(Demand)
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r}
vic_elec %>% gg_season(Demand, period="week")
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r}
vic_elec %>% gg_season(Demand, period="day")
```

## Seasonal subseries plots

  * Data for each season collected together in time plot as separate time series.
  * Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be visualized.
  * In R: `gg_subseries()`

## Quarterly Australian Beer Production

```{r}
beer %>% gg_subseries(Beer)
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism %>%
  filter(Purpose=="Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot, echo=TRUE, dependson="holidays", fig.height=3.9}
holidays %>% autoplot(Trips) +
  ylab("thousands of trips") + xlab("Year") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=4, fig.height=5, out.width="45%"}
holidays %>% gg_season(Trips) +
  ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal subseries plots
\fontsize{9}{10}\sf

```{r graphics2, fig.height=4.1}
holidays %>%
  gg_subseries(Trips) + ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Calendar plots
\fontsize{11}{12}\sf

```{r sugrrants, eval=FALSE}
library(sugrrants)
vic_elec %>%
  filter(year(Date) == 2014) %>%
  mutate(Hour = hour(Time)) %>%
  frame_calendar(x = Hour, y = Demand, date = Date,
    nrow = 4) %>%
  ggplot(aes(x = .Hour, y = .Demand, group = Date)) +
  geom_line() -> p1
prettify(p1, size = 3,
  label.padding = unit(0.15, "lines"))
```

## Calendar plots
\fontsize{10}{11}\sf

```{r sugrrants2, echo=FALSE, fig.height=4.8, fig.width=7.5}
library(sugrrants)
vic_elec %>%
  filter(year(Date) == 2014) %>%
  mutate(Hour = hour(Time)) %>%
  frame_calendar(x = Hour, y = Demand, date = Date, nrow = 4) %>%
  ggplot(aes(x = .Hour, y = .Demand, group = Date)) +
  geom_line() -> p1
prettify(p1, size = 3, label.padding = unit(0.15, "lines"))
```

# Lab Session 3
## Lab Session 3

Look at the quarterly tourism data for the Snowy Mountains

```r
snowy <- filter(tourism,
  Region == "Snowy Mountains")
```

  - Use `autoplot()`, `gg_season()` and `gg_subseries()` to explore the data.
  - What do you learn?


# Decompositions

## Time series decomposition

Trend-Cycle
: aperiodic changes in level over time.

Seasonal
: (almost) periodic changes in level due to seasonal factors (e.g., the quarter of the year, the month, or day of the week).

\begin{block}{Additive decomposition}\vspace*{-0.3cm}
\[ y_t = S_t + T_t + R_t \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}


## STL decomposition

\fontsize{13}{14}\sf

  *  STL: "Seasonal and Trend decomposition using Loess"
  *  Very versatile and robust.
  *  Seasonal component allowed to change over time, and rate of change controlled by user.
  *  Smoothness of trend-cycle also controlled by user.
  *  Optionally robust to outliers
  *  Not trading day or calendar adjustments.
  *  Only additive.
  *  Take logs to get multiplicative decomposition.
  *  Use Box-Cox transformations to get other decompositions.

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot2, echo=TRUE, dependson="holidays", fig.height=3.9}
holidays %>% autoplot(Trips) +
  ylab("thousands of trips") + xlab("Year") +
  ggtitle("Australian domestic holiday nights")
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r stlagain2, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4.3}
holidays %>%
  STL(Trips ~ season(window="periodic"), robust=TRUE) %>%
  autoplot()
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r stlagain, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4.3}
holidays %>%
  STL(Trips ~ season(window = 5), robust = TRUE) %>%
  autoplot()
```

## STL decomposition
\fontsize{11}{13}\sf

```r
holidays %>%
  STL(Trips ~ trend(window=15) + season(window=13),
      robust = TRUE)
```

\fontsize{14}{16}\sf

  *  `trend(window = ?)` controls wiggliness of trend component.
  *  `season(window = ?)` controls variation on seasonal component.
  * `STL()` chooses `season(window=13)` by default
  * A large seasonal window is equivalent to setting `window="periodic"`.
  * Odd numbers should be used for symmetry.

## Holidays decomposition
\fontsize{9}{10}\sf

```{r dable}
dcmp <- holidays %>% STL(Trips)
dcmp
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays3, fig.height=4.6}
dcmp %>% gg_subseries(season_year)
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays-trend, message=FALSE, warning=FALSE, fig.height=4.3}
autoplot(dcmp, trend, scale_bars=FALSE) +
  autolayer(holidays, alpha=0.4)
```

# Lab Session 4

## Lab Session 4

Repeat the decomposition using

\fontsize{11}{14}\sf

```r
holidays %>%
  STL(Trips ~ season(window=7) + trend(window=11)) %>%
  autoplot()
```

\fontsize{14}{15}\sf
What happens as you change `season(window = ???)` and `trend(window = ???)`?

