---
title: "Tidy Time Series & Forecasting in R"
author: "2. Time series graphics"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
```

# Seasonal plots

## Seasonal plots

  * Data plotted against the individual "seasons" in which the data were observed.  (In this case a "season" is a month.)
  * Something like a time plot except that the data from each season are overlapped.
  * Enables the underlying seasonal pattern to be seen more clearly, and also allows any substantial departures from the seasonal pattern to be easily identified.
  * In R: `gg_season()`

## Quarterly Australian Beer Production
\fontsize{13}{15}\sf

```{r, fig.height=3}
beer <- aus_production %>%
  select(Quarter, Beer) %>%
  filter(year(Quarter) >= 1992)
beer %>% autoplot(Beer)
```

## Quarterly Australian Beer Production

```{r}
beer %>% gg_season(Beer, labels = "right")
```

## Multiple seasonal periods
\fontsize{10}{11}\sf

```{r}
vic_elec
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand)
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "week")
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "day")
```

## Seasonal subseries plots

  * Data for each season collected together in time plot as separate time series.
  * Enables the underlying seasonal pattern to be seen clearly, and changes in seasonality over time to be visualized.
  * In R: `gg_subseries()`

## Quarterly Australian Beer Production

```{r}
beer %>% gg_subseries(Beer)
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot, echo=TRUE, dependson="holidays", fig.height=3.9}
holidays %>% autoplot(Trips) +
  ylab("thousands of trips") + xlab("Year") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=4, fig.height=5, out.width="45%"}
holidays %>% gg_season(Trips) +
  ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal subseries plots
\fontsize{9}{10}\sf

```{r graphics2, fig.height=4.1}
holidays %>%
  gg_subseries(Trips) + ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Calendar plots
\fontsize{9}{10}\sf

```{r sugrrants, eval=FALSE}
library(sugrrants)
vic_elec %>%
  filter(year(Date) == 2014) %>%
  mutate(Hour = hour(Time)) %>%
  frame_calendar(
    x = Hour, y = Demand, date = Date,
    nrow = 4
  ) %>%
  ggplot(aes(x = .Hour, y = .Demand, group = Date)) +
  geom_line() -> p1
prettify(p1,
  size = 3,
  label.padding = unit(0.15, "lines")
)
```

\fontsize{12}{14}\sf

`frame_calendar()` makes a compact calendar plot, `facet_calendar()` provides an easier ggplot2 integration.

## Calendar plots
\fontsize{10}{11}\sf

```{r sugrrants2, ref.label="sugrrants", echo=FALSE, fig.height=4.8, fig.width=7.5}

```

# Lab Session 3
## Lab Session 3

1. Look at the quarterly tourism data for the Snowy Mountains

    ```r
    snowy <- tourism %>%
      filter(Region == "Snowy Mountains")
    ```

    - Use `autoplot()`, `gg_season()` and `gg_subseries()` to explore the data.
    - What do you learn?

2. Produce a calendar plot for the `pedestrian` data from one location and one year.

# Seasonal or cyclic?

## Time series patterns

Trend
  : pattern exists when there is a long-term increase or decrease in the data.

Seasonal
  : pattern exists when a series is influenced by seasonal factors (e.g., the quarter of the year, the month, or day of the week).

Cyclic
  : pattern exists when data exhibit rises and falls that are \emph{not of fixed period} (duration usually of at least 2 years).

## Time series components

### Differences between seasonal and cyclic patterns:

* seasonal pattern constant length; cyclic pattern variable length
* average length of cycle longer than length of seasonal pattern
* magnitude of cycle more variable than magnitude of seasonal pattern

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
aus_production %>%
  filter(year(Quarter) >= 1980) %>%
  autoplot(Electricity) + ylab("GWh") +
  ggtitle("Australian electricity production")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, warning=FALSE, fig.height=3}
aus_production %>%
  autoplot(Bricks) +
  ggtitle("Australian clay brick production") +
  xlab("Year") + ylab("million units")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
us_employment %>%
  filter(Title == "Retail Trade", year(Month) >= 1980) %>%
  autoplot(Employed / 1e3) +
  ggtitle("Retail employment, USA") + ylab("Million people")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
gafa_stock %>%
  filter(Symbol == "AMZN", year(Date) >= 2018) %>%
  autoplot(Close) +
  ggtitle("Amazon closing stock price") +
  xlab("Day") + ylab("$")
```

## Time series patterns
\fontsize{10}{10}\sf

```{r, fig.height=3}
pelt %>%
  autoplot(Lynx) +
  ggtitle("Annual Canadian Lynx Trappings") +
  xlab("Year") + ylab("Number trapped")
```

## Seasonal or cyclic?

\alert{Differences between seasonal and cyclic patterns:}

  * seasonal pattern constant length; cyclic pattern variable length
  * average length of cycle longer than length of seasonal pattern
  * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

\begin{alertblock}{}
The timing of peaks and troughs is predictable with seasonal data, but unpredictable in the long term with cyclic data.
\end{alertblock}

# Lag plots and autocorrelation

## Example: Beer production
\fontsize{11}{12}\sf

```{r}
new_production <- aus_production %>%
  filter(year(Quarter) >= 1992)
new_production
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6.5, fig.width=6.5, out.width="8cm"}
new_production %>% gg_lag(Beer)
```

## Example: Beer production
\fontsize{13}{15}\sf

```{r, fig.height=6.5, fig.width=6.5, out.width="8cm"}
new_production %>% gg_lag(Beer, geom = "point")
```

## Lagged scatterplots

  * Each graph shows $y_t$ plotted against $y_{t-k}$ for
different values of $k$.
  * The autocorrelations are the correlations associated
with these scatterplots.
  * ACF (autocorrelation function):
     * $r_1=\text{Correlation}(y_{t}, y_{t-1})$
     * $r_2=\text{Correlation}(y_{t}, y_{t-2})$
     * $r_3=\text{Correlation}(y_{t}, y_{t-3})$
     * etc.
  * If there is **seasonality**, the ACF at the seasonal lag (e.g., 12 for monthly data) will be **large and positive**.

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r, echo=TRUE}
new_production %>% ACF(Beer, lag_max = 9)
```

## Autocorrelation

Results for first 9 lags for beer data:

\fontsize{11}{13}\sf

```{r beeracf, fig.height=2.5}
new_production %>%
  ACF(Beer, lag_max = 9) %>%
  autoplot()
```

\vspace*{10cm}

## ACF

```{r, fig.height=3, echo=TRUE}
new_production %>%
  ACF(Beer) %>%
  autoplot()
```

## Australian holidays
\fontsize{9}{10}\sf

```{r tourismacf}
holidays %>% ACF(Trips)
```

## Australian holidays
\fontsize{9}{10}\sf

```{r tourismacf2, fig.height=6}
holidays %>%
  ACF(Trips) %>%
  autoplot()
```

## Trend and seasonality in ACF plots

- When data have a trend, the autocorrelations for small lags tend to be large and positive.
- When data are seasonal, the autocorrelations will be larger at the seasonal lags (i.e., at multiples of the seasonal frequency)
- When data are trended and seasonal, you see a combination of these effects.

## US retail trade employment
\fontsize{11}{12}\sf

```{r}
retail <- us_employment %>%
  filter(Title == "Retail Trade", year(Month) >= 1980)
retail %>% autoplot(Employed)
```

## US retail trade employment
\fontsize{11}{12}\sf

```{r}
retail %>%
  ACF(Employed, lag_max = 48) %>%
  autoplot()
```

## Google stock price
\fontsize{12}{14}\sf

```{r google-2015}
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) == 2015) %>%
  select(Date, Close)
google_2015
```

## Google stock price

```{r}
google_2015 %>% autoplot(Close)
```

## Google stock price
\fontsize{12}{16}\sf

```r
google_2015 %>%
  ACF(Close, lag_max=100)
# Error: Can't handle tsibble of irregular interval.
```

\pause

```{r}
google_2015
```

## Google stock price
\fontsize{11}{13}\sf

```{r}
google_2015 <- google_2015 %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index = trading_day, regular = TRUE)
google_2015
```

## Google stock price

```{r}
google_2015 %>%
  ACF(Close, lag_max = 100) %>%
  autoplot()
```

# Lab Session 4

## Lab Session 4

We have introduced the following functions: `gg_lag` and `ACF`. Use these functions to explore the four time series: `Bricks` from `aus_production`, `Lynx` from `pelt`, `Close` price of Amazon from `gafa_stock`, `Demand` from `vic_elec`. Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

## Which is which?

```{r, fig.height=6, fig.width=12, echo=FALSE, warning=FALSE, out.width="11.5cm"}
cowtemp <- as_tsibble(fma::cowtemp)
USAccDeaths <- as_tsibble(USAccDeaths)
AirPassengers <- as_tsibble(AirPassengers)
mink <- as_tsibble(fma::mink)
tp1 <- autoplot(cowtemp, value) + xlab("") + ylab("chirps per minute") +
  ggtitle("1. Daily temperature of cow")
tp2 <- autoplot(USAccDeaths, value) + xlab("") + ylab("thousands") +
  ggtitle("2. Monthly accidental deaths")
tp3 <- autoplot(AirPassengers, value) + xlab("") + ylab("thousands") +
  ggtitle("3. Monthly air passengers")
tp4 <- autoplot(mink, value) + xlab("") + ylab("thousands") +
  ggtitle("4. Annual mink trappings")
acfb <- ACF(cowtemp, value) %>% autoplot() + xlab("") + ggtitle("B") + ylim(-0.5, 1)
acfa <- ACF(USAccDeaths, value) %>% autoplot() + xlab("") + ggtitle("A") + ylim(-0.5, 1)
acfd <- ACF(AirPassengers, value) %>% autoplot() + xlab("") + ggtitle("D") + ylim(-0.5, 1)
acfc <- ACF(mink, value) %>% autoplot() + xlab("") + ggtitle("C") + ylim(-0.5, 1)
gridExtra::grid.arrange(tp1, tp2, tp3, tp4,
  acfa, acfb, acfc, acfd,
  nrow = 2
)
```

# White noise

## Example: White noise
\fontsize{11}{13}\sf

```{r, fig.height=3.3}
wn <- tsibble(t = seq(36), y = rnorm(36), index = t)
wn %>% autoplot(y)
```

\only<2>{
\begin{textblock}{10}(1.4,7.2)\fontsize{13}{15}\sf
\begin{alertblock}{}
White noise data is uncorrelated across time with zero mean and constant variance.

(Technically, we require independence as well.)
\end{alertblock}
\end{textblock}}

\vspace*{10cm}

## Example: White noise
\fontsize{11}{11}\sf

```r
wn %>% ACF(y)
```

\fontsize{10}{11}\sf\tabcolsep=0.1cm

```{r, echo=FALSE}
wn %>%
  ACF(y, lag_max = 10) %>%
  as_tibble() %>%
  pivot_wider(names_from = lag, values_from = acf) %>%
  rename_all(function(x) {
    paste("$r_{", x, "}$", sep = "")
  }) %>%
  knitr::kable(
    booktabs = TRUE,
    escape = FALSE, align = "c", digits = 3,
    format.args = list(nsmall = 3)
  )
```

```{r, echo=FALSE, fig.height=2}
wn %>%
  ACF(y) %>%
  autoplot()
```

\pause\fontsize{11}{12}\sf

 * Sample autocorrelations for white noise series.
 * Expect each autocorrelation to be close to zero.
 * Blue lines show 95% critical values.

\vspace*{10cm}

## Example: Pigs slaughtered

\fontsize{11}{12}\sf

```{r, fig.height=3}
pigs <- aus_livestock %>%
  filter(State == "Victoria", Animal == "Pigs",
         year(Month) >= 2014)
pigs %>% autoplot(Count / 1e3) +
  xlab("Year") + ylab("Thousands") +
  ggtitle("Number of pigs slaughtered in Victoria")
```

## Example: Pigs slaughtered

```{r}
pigs %>%
  ACF(Count) %>%
  autoplot()
```

## Example: Pigs slaughtered

Monthly total number of pigs slaughtered
in the state of Victoria, Australia, from January 2014 through December 2018
(Source: Australian Bureau of Statistics.)\pause

  * Difficult to detect pattern in time plot.
  * ACF shows significant autocorrelation for lag 2 and 12.
  * Indicate some slight seasonality.

\pause

These show the series is **not a white noise series**.

# Lab Session 5
## Lab Session 5

You can compute the daily changes in the Google stock price in 2018 using

\fontsize{10.5}{13}\sf

```{r, eval = FALSE}
dgoog <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2018) %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index = trading_day, regular = TRUE) %>%
  mutate(diff = difference(Close))
```

\fontsize{14}{16}\sf

Does `diff` look like white noise?
