---
title: "Tidy Time Series & Forecasting in R"
date: "robjhyndman.com/workshop2020"
author: "4. Seasonality and trends"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    keep_tex: no
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 60)
library(fpp3)
library(purrr)
```


# Time series decompositions

## Time series decomposition

Trend-Cycle
: aperiodic changes in level over time.

Seasonal
: (almost) periodic changes in level due to seasonal factors (e.g., the quarter of the year, the month, or day of the week).

\begin{block}{Additive decomposition}\vspace*{-0.3cm}
\[ y_t = S_t + T_t + R_t \]
\end{block}
\begin{tabular}{@{}llp{8cm}@{}}
where & $y_t=$ & data at period $t$ \\
      & $T_t=$ & trend-cycle component at period $t$\\
      & $S_t=$ & seasonal component at period $t$ \\
      & $R_t=$ & remainder component at period $t$
\end{tabular}


## STL decomposition

\fontsize{13}{14}\sf

  *  STL: "Seasonal and Trend decomposition using Loess"
  *  Very versatile and robust.
  *  Seasonal component allowed to change over time, and rate of change controlled by user.
  *  Smoothness of trend-cycle also controlled by user.
  *  Optionally robust to outliers
  *  Not trading day or calendar adjustments.
  *  Only additive.
  *  Take logs to get multiplicative decomposition.
  *  Use Box-Cox transformations to get other decompositions.

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays, include=FALSE}
holidays <- tourism %>%
  filter(Purpose=="Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r holidays-plot2, echo=TRUE, dependson="holidays", fig.height=3.9}
holidays %>% autoplot(Trips) +
  ylab("thousands of trips") + xlab("Year") +
  ggtitle("Australian domestic holiday nights")
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r stlagain2, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4.3}
holidays %>%
  STL(Trips ~ season(window="periodic"), robust=TRUE) %>%
  autoplot()
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r stlagain, echo=TRUE, warning=FALSE, fig.width=8, fig.height=4.3}
holidays %>%
  STL(Trips ~ season(window = 5), robust = TRUE) %>%
  autoplot()
```

## STL decomposition
\fontsize{11}{13}\sf

```r
holidays %>%
  STL(Trips ~ trend(window=15) + season(window=13),
      robust = TRUE)
```

\fontsize{14}{16}\sf

  *  `trend(window = ?)` controls wiggliness of trend component.
  *  `season(window = ?)` controls variation on seasonal component.
  * `STL()` chooses `season(window=13)` by default
  * A large seasonal window is equivalent to setting `window="periodic"`.
  * Odd numbers should be used for symmetry.

## Holidays decomposition
\fontsize{9}{10}\sf

```{r dable}
dcmp <- holidays %>% STL(Trips)
dcmp
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays3, fig.height=4.6}
dcmp %>% gg_subseries(season_year)
```

## Holidays decomposition
\fontsize{9}{10}\sf

```{r holidays-trend, message=FALSE, warning=FALSE, fig.height=4.3}
autoplot(dcmp, trend, scale_bars=FALSE) +
  autolayer(holidays, alpha=0.4)
```

# Lab Session 7

## Lab Session 7

Repeat the decomposition using

\fontsize{11}{14}\sf

```r
holidays %>%
  STL(Trips ~ season(window=7) + trend(window=11)) %>%
  autoplot()
```

\fontsize{14}{15}\sf
What happens as you change `season(window = ???)` and `trend(window = ???)`?

