---
title: "Tidy Time Series & Forecasting in R"
date: "robjhyndman.com/workshop2020"
author: "7. Exponential smoothing"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 11)
)
options(digits = 3, width = 66)
library(tidyverse)
library(fpp3)
usmelec <- as_tsibble(fpp2::usmelec) %>%
  rename(Month = index, Generation = value)
us_change <- readr::read_csv("https://otexts.com/fpp3/extrafiles/us_change.csv")  %>%
  mutate(Time = yearquarter(Time)) %>%
  as_tsibble(index = Time)
eu_retail <- as_tsibble(fpp2::euretail)
h02 <- tsibbledata::PBS %>%
  filter(ATC2 == "H02") %>%
  summarise(Cost = sum(Cost))
melsyd <- tsibbledata::ansett %>%
  filter(Airports == "MEL-SYD")
austa <- as_tsibble(fpp2::austa) %>%
  rename(Year = index, Visitors = value)
```


# Exponential smoothing

## Historical perspective

 * Developed in the 1950s and 1960s as methods (algorithms) to produce point forecasts.
 * Combine a "level", "trend" (slope) and "seasonal" component to describe a time series.
 * The rate of change of the components are controlled by "smoothing parameters":\newline $\alpha$, $\beta$ and $\gamma$ respectively.
  * Need to choose best values for the smoothing parameters (and initial states).
  * Equivalent ETS state space models developed in the 1990s and 2000s.

## A model for levels, trends, and seasonalities
\fontsize{13}{14}\sf

We want a model that captures the level ($\ell_t$), trend ($b_t$) and seasonality ($s_t$).

\alert{How do we combine these elements?}

\pause

\begin{block}{Additively?}
$y_t = \ell_{t-1} + b_{t-1} + s_{t-m} + \varepsilon_t$
\end{block}\pause
\begin{block}{Multiplicatively?}
$y_t = \ell_{t-1}b_{t-1}s_{t-m}(1 + \varepsilon_t)$
\end{block}\pause
\begin{block}{Perhaps a mix of both?}
$y_t = (\ell_{t-1} + b_{t-1}) s_{t-m} + \varepsilon_t$
\end{block}\pause

\alert{How do the level, trend and seasonal components evolve over time?}

## ETS models

\begin{block}{}
\hspace*{-0.25cm}\begin{tabular}{l@{}p{2.3cm}@{}c@{}l}
\structure{General n\rlap{otation}}
    &       & ~E T S~  & ~:\hspace*{0.3cm}\textbf{E}xponen\textbf{T}ial \textbf{S}moothing               \\ [-0.2cm]
    & \hfill{$\nearrow$\hspace*{-0.1cm}}        & {$\uparrow$} & {\hspace*{-0.2cm}$\nwarrow$} \\
    & \hfill{\textbf{E}rror\hspace*{0.2cm}} & {\textbf{T}rend}      & {\hspace*{0.2cm}\textbf{S}eason}
\end{tabular}
\end{block}

\alert{\textbf{E}rror:} Additive (`"A"`) or multiplicative (`"M"`)
\pause

\alert{\textbf{T}rend:} None (`"N"`), additive (`"A"`), multiplicative (`"M"`), or damped (`"Ad"` or `"Md"`).
\pause

\alert{\textbf{S}easonality:} None (`"N"`), additive (`"A"`) or multiplicative (`"M"`)

## ETS(A,N,N): SES with additive errors

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation}&& \hat{y}_{T+h|T} &= \ell_T \\
\text{Measurement equation}&& y_t &= \ell_{t-1} + \varepsilon_t\\
\text{State equation}&& \ell_t&=\ell_{t-1}+\alpha \varepsilon_t
\end{align*}
\end{block}
where $\varepsilon_t\sim\text{NID}(0,\sigma^2)$.
\pause

  * "innovations" or "single source of error" because equations have the same error process, $\varepsilon_t$.
  * Measurement equation: relationship between observations and states.
  * Transition/state equation(s): evolution of the state(s) through time.

\vspace*{10cm}

## ETS(M,N,N): SES with multiplicative errors

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation}&& \hat{y}_{T+h|T} &= \ell_T \\
\text{Measurement equation}&& y_t &= \ell_{t-1}(1 + \varepsilon_t)\\
\text{State equation}&& \ell_t&=\ell_{t-1}(1+\alpha \varepsilon_t)
\end{align*}
\end{block}
where $\varepsilon_t\sim\text{NID}(0,\sigma^2)$.
\pause

  * Models with additive and multiplicative errors with the same parameters generate the same point forecasts but different prediction intervals.

\vspace*{10cm}

# Trend methods

## Holt's linear trend

\begin{block}{Additive errors: ETS(A,A,N)}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation}&& \hat{y}_{T+h|T} &= \ell_T + hb_T\\
\text{Measurement equation}&& y_t &= \ell_{t-1}+b_{t-1}+\varepsilon_t\\
\text{State equations}&&       \ell_t&=\ell_{t-1}+b_{t-1}+\alpha \varepsilon_t\\
&&      b_t&=b_{t-1}+\beta \varepsilon_t
\end{align*}
\end{block}
\pause

\begin{block}{Multiplicative errors: ETS(M,A,N)}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation}&& \hat{y}_{T+h|T} &= \ell_T + hb_T\\
\text{Measurement equation}&& y_t &= (\ell_{t-1}+b_{t-1})(1+\varepsilon_t)\\
\text{State equations}&&       \ell_t&=(\ell_{t-1}+b_{t-1})(1+\alpha \varepsilon_t)\\
&&      b_t&=b_{t-1}+\beta \varepsilon_t
\end{align*}
\end{block}

## Example: Australian population

\fontsize{9}{9}\sf

```{r holt-fit, echo=TRUE}
aus_economy <- global_economy %>% filter(Code == "AUS") %>%
  mutate(Pop = Population/1e6)
fit <- aus_economy %>% model(AAN = ETS(Pop))
report(fit)
```

## Example: Australian population

\fontsize{10}{11}\sf

```{r holt-cmp, echo=TRUE, dependson='holt-fit'}
components(fit)
```

## Example: Australian population

\fontsize{10}{11}\sf

```{r holt-cmp-plot, echo=TRUE, dependson='holt-fit', fig.height=4.5}
components(fit) %>% autoplot()
```

## Example: Australian population

\fontsize{12}{12}\sf

```{r holt-fc, echo=TRUE, cache=TRUE, dependson='holt-fit'}
fit %>%
  forecast(h = 20) %>%
  autoplot(aus_economy) +
  ylab("Population") + xlab("Year")
```

## ETS(A,Ad,N): Damped trend method
\fontsize{14}{16}\sf

\begin{block}{Additive errors}\vspace*{-0.2cm}
\begin{align*}
\text{Forecast equation}&& \hat{y}_{T+h|T} &= \ell_T + (\phi + \cdots + \phi^{h-1})b_T\\
\text{Measurement equation}&& y_t &= (\ell_{t-1}+\phi b_{t-1})+\varepsilon_t\\
\text{State equations}&&       \ell_t&=(\ell_{t-1}+\phi b_{t-1})+\alpha \varepsilon_t\\
&&      b_t&=\phi b_{t-1}+\beta \varepsilon_t
\end{align*}
\end{block}
\pause

  * Damping parameter $0<\phi<1$.
  * If $\phi=1$, identical to Holt's linear trend.
  * As $h\rightarrow\infty$, $\pred{y}{T+h}{T}\rightarrow \ell_T+\phi b_T/(1-\phi)$.
  * Short-run forecasts trended, long-run forecasts constant.

## Example: Australian population
\fontsize{12}{12}\sf

```{r, echo=TRUE, fig.height=3.6}
aus_economy %>%
  model(holt = ETS(Pop ~ trend("Ad"))) %>%
  forecast(h = 20) %>%
  autoplot(aus_economy)
```

## Example: National populations

\fontsize{9}{9}\sf

```{r popfit, echo=TRUE, cache=TRUE}
fit <- global_economy %>%
  mutate(Pop = Population/1e6) %>%
  model(ets = ETS(Pop))
fit
```

## Example: National populations
\fontsize{12}{12}\sf

```{r popfc, echo=TRUE, cache=TRUE, dependson="popfit"}
fit %>%
  forecast(h = 5)
```

# Lab Session 14
## Lab Session 14

Consider the data set `eggs`, the price of a dozen eggs in the United States from 1900-1993. Experiment with the various options in the `ETS()`` function to see how much the forecasts change with damped trend, or with a Box-Cox transformation. Try to develop an intuition of what each is doing to the forecasts. How can we stop the forecasts becoming negative?

[Hint: use `h=100` when forecasting, so you can clearly see the differences between the various options when plotting the forecasts.]

# Seasonal methods

## \fontsize{16}{16}\sf\bfseries ETS(A,A,A): Holt-Winters additive method

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation} && \hat{y}_{t+h|t} &= \ell_{t} + hb_{t} + s_{t+h-m(k+1)}\\
\text{Observation equation}&& y_t&=\ell_{t-1}+b_{t-1}+s_{t-m} + \varepsilon_t\\
\text{State equations}&& \ell_t&=\ell_{t-1}+b_{t-1}+\alpha \varepsilon_t\\
&&        b_t&=b_{t-1}+\beta \varepsilon_t \\
&&s_t &= s_{t-m} + \gamma\varepsilon_t
\end{align*}
\end{block}

* $k=$ integer part of $(h-1)/m$.
* $\sum_i s_i \approx 0$.
* Parameters:&nbsp; $0\le \alpha\le 1$,&nbsp; $0\le \beta^*\le 1$,&nbsp; $0\le \gamma\le 1-\alpha$&nbsp;  and $m=$  period of seasonality (e.g. $m=4$ for quarterly data).

## \fontsize{16}{16}\sf\bfseries ETS(M,A,M): Holt-Winters multiplicative method

\begin{block}{}\vspace*{-0.4cm}
\begin{align*}
\text{Forecast equation} && \hat{y}_{t+h|t} &= (\ell_{t} + hb_{t}) s_{t+h-m(k+1)}\\
\text{Observation equation}&& y_t&= (\ell_{t-1}+b_{t-1})s_{t-m}(1 + \varepsilon_t)\\
\text{State equations}&& \ell_t&=(\ell_{t-1}+b_{t-1})(1+\alpha \varepsilon_t)\\
&&        b_t&=b_{t-1}(1+\beta \varepsilon_t) \\
&&s_t &= s_{t-m}(1 + \gamma\varepsilon_t)
\end{align*}
\end{block}

* $k$ is integer part of $(h-1)/m$.
* $\sum_i s_i \approx m$.
* Parameters:&nbsp; $0\le \alpha\le 1$,&nbsp; $0\le \beta^*\le 1$,&nbsp; $0\le \gamma\le 1-\alpha$&nbsp;  and $m=$  period of seasonality (e.g. $m=4$ for quarterly data).

## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-fit, echo=TRUE}
holidays <- tourism %>%
  filter(Purpose == "Holiday")
fit <- holidays %>% model(ets = ETS(Trips))
fit
```


## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-report}
fit %>% filter(Region=="Snowy Mountains") %>% report()
```

## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-components}
fit %>% filter(Region=="Snowy Mountains") %>% components(fit)
```

## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-components-plot, fig.height=4.3}
fit %>% filter(Region=="Snowy Mountains") %>%
  components(fit) %>% autoplot()
```

## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-forecast}
fit %>% forecast()
```

## Example: Australian holiday tourism

\fontsize{9}{10}\sf

```{r ausholidays-forecast-plot}
fit %>% forecast() %>%
  filter(Region=="Snowy Mountains") %>%
  autoplot(holidays) +
    xlab("Year") + ylab("Overnight trips (thousands)")
```

# ETS taxonomy

## Exponential smoothing models
\fontsize{11}{12}\sf

\begin{block}{}
\begin{tabular}{ll|ccc}
  \multicolumn{2}{l}{\alert{\bf Additive Error}} &        \multicolumn{3}{c}{\bf Seasonal Component}         \\
          \multicolumn{2}{c|}{\bf Trend}         &         N         &         A         &         M         \\
        \multicolumn{2}{c|}{\bf Component}       &     ~(None)~      &    (Additive)     & (Multiplicative)  \\ \cline{3-5}
           &                                     &                   &                   &  \\[-0.3cm]
  N        & (None)                              &       A,N,N       &       A,N,A       &    \st{A,N,M}     \\
           &                                     &                   &                   &  \\[-0.3cm]
  A        & (Additive)                          &       A,A,N       &       A,A,A       &    \st{A,A,M}     \\
           &                                     &                   &                   &  \\[-0.3cm]
  A\damped & (Additive damped)                   &   A,A\damped,N    &   A,A\damped,A    & \st{A,A\damped,M}
\end{tabular}
\end{block}

\begin{block}{}
\begin{tabular}{ll|ccc}
  \multicolumn{2}{l}{\alert{\bf Multiplicative Error}} &     \multicolumn{3}{c}{\bf Seasonal Component}      \\
             \multicolumn{2}{c|}{\bf Trend}            &      N       &         A         &        M         \\
           \multicolumn{2}{c|}{\bf Component}          &   ~(None)~   &    (Additive)     & (Multiplicative) \\ \cline{3-5}
           &                                           &              &                   &  \\[-0.3cm]
  N        & (None)                                    &    M,N,N     &       M,N,A       &      M,N,M       \\
           &                                           &              &                   &  \\[-0.3cm]
  A        & (Additive)                                &    M,A,N     &       M,A,A       &      M,A,M       \\
           &                                           &              &                   &  \\[-0.3cm]
  A\damped & (Additive damped)                         & M,A\damped,N &   M,A\damped,A    &   M,A\damped,M
\end{tabular}
\end{block}

## Estimating ETS models

  * Smoothing parameters $\alpha$, $\beta$, $\gamma$ and $\phi$, and the initial states $\ell_0$, $b_0$, $s_0,s_{-1},\dots,s_{-m+1}$ are estimated by maximising the "likelihood" = the probability of the data arising from the specified model.
  * For models with additive errors equivalent to minimising SSE.
  * For models with multiplicative errors, \textbf{not} equivalent to minimising SSE.

## Model selection
\fontsize{13}{15}\sf

\begin{block}{Akaike's Information Criterion}
\[
\text{AIC} = -2\log(\text{L}) + 2k
\]
\end{block}\vspace*{-0.2cm}
where $L$ is the likelihood and $k$ is the number of parameters initial states estimated in the model.\pause

\begin{block}{Corrected AIC}
\[
\text{AIC}_{\text{c}} = \text{AIC} + \frac{2(k+1)(k+2)}{T-k}
\]
\end{block}
which is the AIC corrected (for small sample bias).
\pause
\begin{block}{Bayesian Information Criterion}
\[
\text{BIC} = \text{AIC} + k(\log(T)-2).
\]
\end{block}

## AIC and cross-validation

\Large

\begin{alertblock}{}
Minimizing the AIC assuming Gaussian residuals is asymptotically equivalent to minimizing one-step time series cross validation MSE.
\end{alertblock}

## Automatic forecasting

**From Hyndman et al.\ (IJF, 2002):**

1. Apply each model that is appropriate to the data.
Optimize parameters and initial values using MLE.
1. Select best method using AICc.
1. Produce forecasts using best method.
1. Obtain forecast intervals using underlying state space model.

* Method performed very well in M3 competition.
* Used as a benchmark in the M4 competition.

# Lab Session 15

## Lab Session 15

Find an ETS model for the Gas data from `aus_production`.

  * Why is multiplicative seasonality necessary here?
  * Experiment with making the trend damped.

