---
title: "Tidy Time Series & Forecasting in R"
author: "9. Dynamic regression"
toc: true
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 7
    fig_height: 3.5
    includes:
      in_header: header.tex
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, cache = TRUE, warning = FALSE, message = FALSE,
  dev.args = list(bg = grey(0.9), pointsize = 11)
)
library(fpp3)

vic_elec_daily <- vic_elec %>%
  filter(year(Time) == 2014) %>%
  index_by(Date = date(Time)) %>%
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) %>%
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
```

# Regression with ARIMA errors

## Regression with ARIMA errors
\fontsize{13}{15}\sf

\begin{block}{Regression models}\vspace*{-0.2cm}
\[
  y_t = \beta_0 + \beta_1 x_{1,t} + \dots + \beta_k x_{k,t} + \varepsilon_t,
\]
\end{block}

  * $y_t$ modeled as function of $k$ explanatory variables
  * In regression, we assume that $\varepsilon_t$ is white noise.
\vspace*{0.3cm}
\pause
\begin{alertblock}{RegARIMA model}\vspace*{-0.2cm}
\begin{align*}
  y_t &= \beta_0 + \beta_1 x_{1,t} + \dots + \beta_k x_{k,t} + \eta_t,\\
      \eta_t &\sim \text{ARIMA}
\end{align*}
\end{alertblock}

  * Residuals are from ARIMA model.
  * Estimate model in one step using MLE
  * Select model with lowest AICc value.

## US personal consumption and income
\fontsize{9}{10}\sf

```{r usconsumpdata, echo=TRUE}
us_change
```

## US personal consumption and income

```{r usconsump, fig.height=6, fig.width=8.5}
us_change %>%
  pivot_longer(-Quarter, names_to = "variable", values_to = "value") %>%
  ggplot(aes(y = value, x = Quarter, group = variable)) +
  geom_line() + facet_grid(variable ~ ., scales = "free_y") +
  xlab("Year") + ylab("") +
  ggtitle("Quarterly changes in US consumption and personal income")
```

## US personal consumption and income
\fontsize{9}{10}\sf

```{r usconsump_pairs, fig.height=6, fig.width=8.5, echo=TRUE}
us_change %>% as_tibble() %>% select(-Quarter) %>% GGally::ggpairs()
```

## US personal consumption and income

  * No need for transformations or further differencing.
  *  Increase in income does not necessarily translate into instant increase in consumption (e.g., after the loss of a job, it may take a few months for expenses to be reduced to allow for the new circumstances). We will ignore this for now.

## US personal consumption and income
\fontsize{8}{9}\sf

```{r usconsump2, echo=TRUE, fig.height=3}
fit <- us_change %>%
  model(regarima = ARIMA(Consumption ~ Income + Production + Savings + Unemployment))
report(fit)
```

\fontsize{13}{14}\sf\pause\begin{alertblock}{}
Write down the equations for the fitted model.
\end{alertblock}

## US personal consumption and income
\fontsize{8}{9}\sf

```{r , echo=TRUE, fig.height=3.7, dependson='usconsump2'}
gg_tsresiduals(fit)
```

## US personal consumption and income
\fontsize{8}{9}\sf

```{r , echo=TRUE, fig.height=3.7, dependson='usconsump2'}
augment(fit) %>%
  features(.resid, ljung_box, dof = 6, lag = 12)
```

## US personal consumption and income
\fontsize{8}{8}\sf

```{r usconsump3, echo=TRUE, fig.height=3, dependson='usconsump2'}
us_change_future <- new_data(us_change, 8) %>%
  mutate(Income = tail(us_change$Income,1),
         Production =tail(us_change$Production,1),
         Savings = tail(us_change$Savings,1),
         Unemployment = tail(us_change$Unemployment,1))
forecast(fit, new_data = us_change_future) %>%
  autoplot(us_change) +
  labs(x = "Year", y = "Percentage change",
       title = "Forecasts from regression with ARIMA errors")
```

## Forecasting

  * To forecast a regression model with ARIMA errors, we need to forecast the
regression part of the model and the ARIMA part of the model and combine the
results.
  * Some predictors are known into the future (e.g., time, dummies).
  * Separate forecasting models may be needed for other predictors.
  * Forecast intervals ignore the uncertainty in forecasting the predictors.

## Daily electricity demand
\fontsize{12}{13}\sf

Model daily electricity demand as a function of temperature using quadratic regression with ARMA errors.

\fontsize{8}{9}\sf

```{r, echo=TRUE, fig.height=3}
vic_elec_daily %>%
  ggplot(aes(x = Temperature, y = Demand, colour = Day_Type)) +
  geom_point() +
  labs(x = "Maximum temperature", y = "Electricity demand (GW)")
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo=TRUE, fig.height=3.6}
vic_elec_daily %>%
  pivot_longer(c(Demand, Temperature)) %>%
  ggplot(aes(x = Date, y = value)) + geom_line() +
  facet_grid(vars(name), scales = "free_y")
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r dailymodel, echo=TRUE}
fit <- vic_elec_daily %>%
  model(fit = ARIMA(Demand ~ Temperature + I(Temperature^2) +
    (Day_Type == "Weekday")))
report(fit)
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo=TRUE, dependson='dailymodel'}
augment(fit) %>%
  gg_tsdisplay(.resid, plot_type = "histogram")
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo=TRUE, dependson='dailymodel'}
augment(fit) %>%
  features(.resid, ljung_box, dof = 9, lag = 14)
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo=TRUE, dependson='dailymodel'}
# Forecast one day ahead
vic_next_day <- new_data(vic_elec_daily, 1) %>%
  mutate(Temperature = 26, Day_Type = "Holiday")
forecast(fit, vic_next_day)
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo=TRUE}
vic_elec_future <- new_data(vic_elec_daily, 14) %>%
  mutate(
    Temperature = 26,
    Holiday = c(TRUE, rep(FALSE, 13)),
    Day_Type = case_when(
      Holiday ~ "Holiday",
      wday(Date) %in% 2:6 ~ "Weekday",
      TRUE ~ "Weekend"
  )
)
```

## Daily electricity demand
\fontsize{8}{9}\sf

```{r, echo = TRUE, dependson='dailymodel'}
forecast(fit, vic_elec_future) %>%
  autoplot(vic_elec_daily) + ylab("Electricity demand (GW)")
```

# Lab Session 18
## Lab Session 18
\fontsize{11}{12}\sf

Repeat the daily electricity example, but instead of using a quadratic function of temperature, use a piecewise linear function with the "knot" around 20 degrees Celsius (use predictors `Temperature` & `Temp2`). How can you optimize the choice of knot?

The data can be created as follows.

\vspace*{-0.2cm}\fontsize{9}{9.5}\sf

```{r echo=TRUE, eval=FALSE}
vic_elec_daily <- vic_elec %>%
  filter(year(Time) == 2014) %>%
  index_by(Date = date(Time)) %>%
  summarise(
    Demand = sum(Demand) / 1e3,
    Temperature = max(Temperature),
    Holiday = any(Holiday)
  ) %>%
  mutate(
    Temp2 = I(pmax(Temperature - 20, 0)),
    Day_Type = case_when(
      Holiday ~ "Holiday",
      wday(Date) %in% 2:6 ~ "Weekday",
      TRUE ~ "Weekend"))
```

# Dynamic harmonic regression

## Dynamic harmonic regression

 **Combine Fourier terms with ARIMA errors**

\fontsize{13}{14}\sf

### Advantages
   * it allows any length seasonality;
   * for data with more than one seasonal period, you can include Fourier terms of different frequencies;
   * the seasonal pattern is smooth for small values of $K$ (but more wiggly seasonality can be handled by increasing $K$);
   * the short-term dynamics are easily handled with a simple ARMA error.

### Disadvantages
 * seasonality is assumed to be fixed

## Eating-out expenditure
\fontsize{8}{9}\sf

```{r cafe, echo=TRUE, fig.height=3.5, fig.width=8}
aus_cafe <- aus_retail %>%
  filter(
    Industry == "Cafes, restaurants and takeaway food services",
    year(Month) %in% 2004:2018
  ) %>%
  summarise(Turnover = sum(Turnover))
aus_cafe %>% autoplot(Turnover)
```

## Eating-out expenditure

\fontsize{9}{10}\sf

```{r cafefit, dependson='cafe', fig.height=5, echo=TRUE, results='hide'}
fit <- aus_cafe %>% model(
  `K = 1` = ARIMA(log(Turnover) ~ fourier(K = 1) + PDQ(0, 0, 0)),
  `K = 2` = ARIMA(log(Turnover) ~ fourier(K = 2) + PDQ(0, 0, 0)),
  `K = 3` = ARIMA(log(Turnover) ~ fourier(K = 3) + PDQ(0, 0, 0)),
  `K = 4` = ARIMA(log(Turnover) ~ fourier(K = 4) + PDQ(0, 0, 0)),
  `K = 5` = ARIMA(log(Turnover) ~ fourier(K = 5) + PDQ(0, 0, 0)),
  `K = 6` = ARIMA(log(Turnover) ~ fourier(K = 6) + PDQ(0, 0, 0))
)
glance(fit)
```
```{r, echo = FALSE}
glance(fit) %>%
  select(.model, sigma2, log_lik, AIC, AICc, BIC) %>%
  knitr::kable()
```

## Eating-out expenditure

```{r}
cafe_plot <- function(...) {
  fit %>%
    select(...) %>%
    forecast() %>%
    autoplot(aus_cafe) +
    ggtitle(sprintf("Log transformed %s, fourier(K = %s)", model_sum(select(fit, ...)[[1]][[1]]), deparse(..1))) +
    geom_label(
      aes(x = yearmonth("2007 Jan"), y = 4250, label = paste0("AICc = ", format(AICc))),
      data = glance(select(fit, ...))
    ) +
    geom_line(aes(y = .fitted), colour = "red", augment(select(fit, ...))) +
    ylim(c(1500, 5100))
}
```

```{r cafe1, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 1)
```

## Eating-out expenditure

```{r cafe2, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 2)
```

## Eating-out expenditure

```{r cafe3, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 3)
```

## Eating-out expenditure

```{r cafe4, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 4)
```

## Eating-out expenditure

```{r cafe5, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 5)
```

## Eating-out expenditure

```{r cafe6, dependson='cafe', fig.height=5, echo=FALSE}
cafe_plot(K = 6)
```

## Example: weekly gasoline products
\fontsize{8}{8}\sf
```{r, echo = FALSE}
options(width = 70)
```

```{r gasmodel, echo=TRUE}
gasoline <- as_tsibble(fpp2::gasoline)
fit <- gasoline %>% model(ARIMA(value ~ fourier(K = 13) + PDQ(0, 0, 0)))
report(fit)
```

## Example: weekly gasoline products
\fontsize{11}{12}\sf

```{r gasf, echo=TRUE, fig.height=3.5}
forecast(fit, h = "3 years") %>%
  autoplot(gasoline)
```


# Lab Session 19
## Lab Session 19

Repeat Lab Session 18 but using all available data, and handling the annual seasonality using Fourier terms.

# Lagged predictors

## Lagged predictors

Sometimes a change in $x_t$ does not affect $y_t$ instantaneously

\begin{block}{}
\begin{itemize}
  \item $y_t=$ sales, $x_t=$ advertising.
  \item $y_t=$ stream flow, $x_t=$ rainfall.
  \item $y_t=$ size of herd, $x_t=$ breeding stock.
\end{itemize}
\end{block}
\pause

  * These are dynamic systems with input ($x_t$) and output $(y_t)$.
  * $x_t$ is often a leading indicator.
  * There can be multiple predictors.

## Lagged predictors

The model include present and past values of predictor: $x_t,x_{t-1},x_{t-2},\dots.$
\begin{block}{}
\centerline{$
y_t = a + \nu_0x_t + \nu_1x_{t-1} + \dots + \nu_kx_{t-k} + \eta_t$}
\end{block}
where $\eta_t$ is an ARIMA process.\pause

  * $x$ can influence $y$, but $y$ is not allowed to influence $x$.

## Example: Insurance quotes and TV adverts

```{r tvadvertdata}
insurance <- as_tsibble(fpp2::insurance, pivot_longer = FALSE) %>%
  rename(Month = index)
```

```{r tvadvertdata2}
insurance
```

## Example: Insurance quotes and TV adverts

```{r tvadvert, dependson='tvadvertdata'}
insurance %>%
  pivot_longer(c(Quotes, TV.advert)) %>%
  ggplot(aes(x = Month, y = value)) + geom_line() +
  facet_grid(vars(name), scales = "free_y") +
  labs(x = "Year", y = NULL, title = "Insurance advertising and quotations")
```

## Example: Insurance quotes and TV adverts

```{r tvadvertpairs, dependson='tvadvertdata'}
insurance %>%
  mutate(
    lag1 = lag(TV.advert),
    lag2 = lag(lag1)
  ) %>%
  as_tibble() %>%
  select(-Month) %>%
  rename(lag0 = TV.advert) %>%
  pivot_longer(-Quotes, names_to="Lag", values_to="TV_advert") %>%
  ggplot(aes(x = TV_advert, y = Quotes)) + geom_point() +
  facet_grid(. ~ Lag) +
  labs(title = "Insurance advertising and quotations")
```


## Example: Insurance quotes and TV adverts
\fontsize{10}{10}\sf

```{r, echo=TRUE}
fit <- insurance %>%
  # Restrict data so models use same fitting period
  mutate(Quotes = c(NA, NA, NA, Quotes[4:40])) %>%
  model(
    ARIMA(Quotes ~ pdq(d = 0) + TV.advert),
    ARIMA(Quotes ~ pdq(d = 0) + TV.advert + 
                                lag(TV.advert)),
    ARIMA(Quotes ~ pdq(d = 0) + TV.advert + 
                                lag(TV.advert) +
                                lag(TV.advert, 2)),
    ARIMA(Quotes ~ pdq(d = 0) + TV.advert + 
                                lag(TV.advert) +
                                lag(TV.advert, 2) + 
                                lag(TV.advert, 3))
  )
```

## Example: Insurance quotes and TV adverts
\fontsize{10}{10}\sf

```{r, echo=TRUE, results = 'hide'}
glance(fit)
```
```{r, echo = FALSE}
glance(fit) %>%
  transmute(`Lag order` = 0:3, sigma2, log_lik, AIC, AICc, BIC) %>%
  knitr::kable()
```

## Example: Insurance quotes and TV adverts

\fontsize{8}{9}\sf

```{r tvadvertagain, echo=TRUE}
# Re-fit to all data
fit <- insurance %>%
  model(ARIMA(Quotes ~ TV.advert + lag(TV.advert) + pdq(d = 0)))
report(fit)
```

\pause
\vspace{-1em}

```{r tvadvertparam, echo=FALSE, dependson='tvadvertagain'}
# Store coefficients
coef <- rlang::set_names(tidy(fit)$estimate, tidy(fit)$term)
phi1 <- coef["ar1"]
theta1 <- coef["ma1"]
theta2 <- coef["ma2"]
intercept <- coef["intercept"]
gamma0 <- coef["TV.advert"]
gamma1 <- coef["lag(TV.advert)"]
```

###
\begin{align*}
  y_t &= `r format(intercept, digits=3)` +
         `r format(gamma0, digits=3)` x_t +
         `r format(gamma1, digits=2)` x_{t-1} + \eta_t,\\
  \eta_t &= `r format(phi1, digits=3)` \eta_{t-1} +
        \varepsilon_t +
        `r format(theta1, digits=2)` \varepsilon_{t-1} +
        `r format(theta2, digits=2)` \varepsilon_{t-2}.
\end{align*}

## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf

```{r, echo=TRUE, fig.height=3.3}
advert_a <- new_data(insurance, 20) %>%
  mutate(TV.advert = 10)
forecast(fit, advert_a) %>% autoplot(insurance)
```

## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf

```{r, echo=TRUE, fig.height=3.3}
advert_b <- new_data(insurance, 20) %>%
  mutate(TV.advert = 8)
forecast(fit, advert_b) %>% autoplot(insurance)
```

## Example: Insurance quotes and TV adverts
\fontsize{11}{13}\sf

```{r, echo=TRUE, fig.height=3.3}
advert_c <- new_data(insurance, 20) %>%
  mutate(TV.advert = 6)
forecast(fit, advert_c) %>% autoplot(insurance)
```
